### 2019-10-17 doc/Casey1.Rmd
added manuscript current draft, set up directory structure

---
title: "Distinct communities of snow algae bloom at different elevations"
author:
  - name: Casey B. Engstrom
    institute: SFU
  - name: Kurt M. Yakimovich
    institute: SFU
  - name: Lynne M. Quarmby
    institute: SFU
    email: quarmby@sfu.ca
institute:
  - SFU: Dept. Molecular Biology and Biochemistry, Simon Fraser University, 8888 University Drive, Burnaby, BC, V5A 1S6, Canada
date: "`r format(Sys.time())`"
output: 
  word_document:
    # toc: yes
    pandoc_args:
      - '--lua-filter=lua-filters-master/scholarly-metadata.lua'
      - '--lua-filter=lua-filters-master/author-info-blocks.lua'
    reference_docx: /home/cengstro/projects/sa_rbcl/Casey1/tmp/word-template.docx
  pdf_document: 
    number_sections: yes
    # toc: yes
    # toc_depth: 4
    pandoc_args:
      - '--lua-filter=lua-filters-master/scholarly-metadata.lua'
      - '--lua-filter=lua-filters-master/author-info-blocks.lua'
bibliography: /home/cengstro/ownCloud/mendeley_library/bibtex_by_grp/casey1.bib
abstract: |
  *Abstract*: Snow algae grow in melting summer snow across polar and alpine regions, forming green or red colored blooms on the snow surface which can reduce snow surface albedo and increases snow melt rate. Despite over a century of study, the we know little of the diversity of algae species that cause snow algae blooms. We used light microscopy and high-throughput amplicon sequencing of the plastid *rbc*L gene and 18S rDNA to survey snow algae blooms in the Coast Range of British Columbia, Canada. We found distinct communities at different elevations. Our findings suggest that some species of snow algae are widely dispersed, but environmental conditions likely select which snow algae species will proliferate.

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>
<br>
<br>
<!-- Our findings highlight the need to consider the species composition of snow algae blooms in future research of this poorly understood microbiome. -->

### *Keywords*: snow algae rbcL 18S amplicon microbiome alpine community

### *Run title*: Elevation patterns in snow algae communities

### *To be submitted to*: 1) Frontiers in Micb 2) Env Micb 3) Microbial Ecology

<!-- what we now know: (Old studies suggested... here we are blowing open the door. ) -->
<!-- - there are different microbiomes -->
<!-- - alpine vs subalpine -->
<!-- - communities are not distinct, but a gradient!! -->

<!-- completely novel -->
<!-- - chloromonas krienitzii blooms in snow runnels (more seen below treeline), green below, orange above -->
<!-- - chlainomonas is widespread in all blooms -->
<!-- - muramotoi in North America? (if this is a reliable genetic marker) -->
<!-- - Some unknown Chloromonas OTUs with no match in ref databases -->
<!-- - distinct clade Chlainomonas above treeline -->
<!-- - sphere krienitzii, oval, same? or closely related... -->
<!-- - Chlainomonas more at surface than depth -->

<!-- 18S data -->
<!-- write discussion -->

##### pagebreak

# Introduction {-}

Snow algae are cold-adapted microalgae that bloom in melting snow [@Leya2013]. Red pigments produced by some species of snow algae can color vast areas of polar and alpine snowfields, which decreases snow surface albedo, by one estimate enough to substantially increase the rate of snow melt [@Ganey2017]. Loss of snow cover can contribute to drought [@Staudinger2014;@Barnhart2016] and exposes darker surfaces upon meltout, contributing to snow-albedo feedback [@Dery2007]. Thus, snow algae could play an important role in the cryosphere, yet little is known about what species of algae are present, or differences in community composition between snow algae blooms.

Until recently, snow algae species were defined and identified on the basis of morphology. Morphology can be deceptive when identifying microalgae, as the same species can look completely different from one life stage to the next [@Shah2016;@Salome2019]. In the lab, cultures of *Chloromonas* are all morphologically indistinct green biflagellates, but field samples of this genus are dominated by cells with yellow or orange pigment and have thick cell walls with flanges or spikes [@Matsuzaki2019;@Prochazkova2019a]. Cells formerly defined by morphological similarity can be genetically distinct, for example cell types referred to as *Chloromonas cf. nivalis* are now split into multiple species [@Matsuzaki2019;@Prochazhova2019a].
<!-- based on this paragraph, I would not trust the morphology categories in my thing. Ill take em out anyway, but... chlainomonas is distinct, chloro distinct -->

<!-- get reader up to date on snow algae taxonomy... two out of three are poorly explored -->
To date, genetic studies have identified three genera of algae that cause most snow algae blooms: *Chloromonas*, *Sanguina*, and *Chlainomonas*. Only *Chloromonas* contains cultured representatives, with _ named species on GenBank. *Chloromonas* are not known to produce the red pigment. *Sanguina* was recently designated, and contains just two species, one of which *Sanguina nivaloides* is found in alpine and polar snowfields worldwide, and is likely synonymous with some of *Chlamydomonas cf. nivalis* [@Prochazkova2018c]. *Chlainomonas* are less studied, but appear in field samples as large (up to 40 $\mu$m diameter) quadriflagellate oval red cells. In spite of their striking morphological differences, *rbc*L and 18S phylogeny places them within the *Chloromonas* clade.

<!-- how communities vary from patch to patch. -->

<!-- Much of the recent snow algae literature has focused on the eye-catching red snow prevalent throughout polar and alpine regions [@Segawa2018;@Leya2013].  -->

The habitat distribution of snow algae are not well established. Amplicon high-throughput sequencing (HTS) shows that red snow from sites across the Arctic are dominated by reads with high similarity to *Sanguina*, with low relative abundance of *Chloromonas polyptera*, *Chloromonas cf. nivalis*, and *Raphidonema* [@Lutz2016a]. @Ganey2018 suggests that the red pigment of snow algae may be a special adaptation to drier snow, while green snow is restricted to moister areas, although red snow caused by *Chlainomonas* is only reported from waterlogged snow overlying alpine lakes [@Novis2008;@Prochazkova2018c]. Green snow blooms dominated by *Chloromonas* are reported from forested sites and shaded snowbanks [@Terashima2017;@Remias2013a], although some *Chloromonas* may be specially adapted for high light intensity in the alpine [@Prochazkova2019a]. In the Arctic, one green snow patch was dominated by *Microglena* [@Lutz2015]. Microscopy suggests differences in the snow algae community above and below treeline: two studies found red snow containing *Sanguina*-like cells to be prevalent in the alpine zone above treeline, while green and orange snows caused by *Chloromonas cf. brevispina* and *Chloromonas cf. nivalis* were prevalent in the subalpine zone below treeline [@Hoham1979a;@Nedbalova2008a].

<!-- clear objectives section -->
 Whether snow algae blooms are typically dominated by a single species, or contain a diversity of species remains an open question, as high-resolution genetic comparison between communities is lacking.

Thus far, efforts to characterize snow algae communities by amplicon high-throughput sequencing (HTS) have been limited by ambiguous or nonexistent reference sequences and high 18S similarity between snow algae species [@Lutz2018]. To robustly compare snow algae community comparison side-by-side we used rbcL and 18S amplicon HTS and microscopy to survey 35 snow algae communities in our local mountains, to see what species were present, to compare community composition between samples, and see how they are distributed across the landscape.

<!-- These blooms are often dominated by red spherical cells, which until recently were referred to in the literature as *Chlamydomonas cf. nivalis*, but recent taxonomy classifies as *Sanguina nivaloides*, although synonymy was not established [@Prochazkova2019].  -->
<!-- What species? Are snow algae communities distinct, or overlapping gradient?  -->

<!-- Snow algae diversity is poorly understood, and whether different species inhabit different niches remains an open question.  -->

<!-- A diversity of algae species and cell morphologies are reported from snow algae blooms, yet little is known about the habitat distribution of snow algae species. -->
<!-- why is this interesting? -->
<!-- make it obvious what the paper is about early as possible -->

<!-- yet community composition is perceived to be relatively homogeneous between sites [@Lutz2016a]; whether there may be different types of snow algae blooms has not been adequately explored. -->

<!-- but their impact on alpine food webs and nutrient cycling has not been explored. -->

<!-- Until recently, snow algae species were defined on the basis of cell morphology. -->

<!-- *Sanguina* currently has only two named species, , and . The competitive exclusion principle suggests that within the melting snowpack microbiome there exist different niches, for example different snow moisture content, shade, or nutrient input from trees [@Hoham1976].  -->

<!-- Chloro polyptera absent in rbcL. could we detect this in our samples by 18S? how similar is it to other Chloromonas? -->

<!-- Red cells of *Chlamydomonas cf. nivalis* can be either cosmopolitan or endemic between the poles [@Segawa2018]. -->

<!-- What species of Chloromonas? (Chlainomonas within)  -->

<!-- What potentially novel species look similar, but are genetically distinct? -->

<!-- Are all snow algae communities the same, or different?  Do the differences in cell morphology correspond to genetically different species? What described species are found in the snow? Which of these may be unknown species that looked similar in previous surveys?  -->
<!-- What do we now know? Chlainomonas hi low, above and below treeline specific otus, Chloro krienitzii is widespread in runnels, esp below treeline, green below turning orange on surface. Sanguina less common below treeline.  -->

<!-- what we now know: (Old studies suggested... here we are blowing open the door. ) -->

<!-- - gradient, dom morph vs mix -->
<!--   - chlainomonas is widespread in all blooms -->
<!--   - X17 -->

<!-- - alpine vs subalpine -->
<!--   - Some unknown Chloromonas OTUs with no match in ref databases -->
<!--   - distinct clade Chlainomonas above treeline -->

<!-- - krienitzii -->
<!--   - in snow runnels (more seen below treeline), green below, orange above -->
<!--   - sphere krienitzii, oval, same? or closely related... -->
<!--   - Chlainomonas more at surface than depth -->

# Results {-}

**A gradient between snow algae communities**. Variation in snow algae communities was primarily driven by the presence or absence of *Sanguina nivaloides*, *Chloromonas krienitzii*, *Chlainomonas rubra*, or *Chlainomonas* OTU3 (Fig. 2a). <!--to use named OTUs, or not. The line drawn by our classifier is also arbitrary, depending on database etc but less so than me circling some dots. But I could make a definition in methods, refer to OTU by the most abundant species in it--> However, most samples contained a mix of species; few samples were relatively "pure". However, sample comparison with weighted UniFrac [@Lozupone2005] found little difference between *Chlainomonas* and *Chloromonas* dominated sites, and the presence or absence of *Sanguina* was the driving factor, which was generally found in higher relative abundance at higher elevation (Fig. 2b).

Although *Chlainomonas* is reported to only grow in slushy snow above alpine tarns [@Novis2008;@Prochazkova2018c], we found *Chlainomonas* was present in low relative abundance at most sites sampled. Only one site high in Chlainomonas was near water (S9), the other Chlainomonas dominated sites were not near water, nor was the snow noticeably mushier than the surrounding snow. This does not rule out the possibility that these sites contained high moisture at some point; further sampling is needed to determine its habitat distribution.

<!-- One *Chlainomonas* dominated sample (S9) was from snow near the edge of a pool, but we also collected *Chlainomonas* at sites where we did not see evidence of water nearby (S13, G16.1), and one from a pink snow runnel (S6) that looked the same as the runnels dominated by *C. krienitzii*. *Chlainomonas* dominated samples appeared indistinguishable to pink snow caused by *Sanguina* to the naked eye, although it was less frequently collected than *Sanguina*. This suggests that *Chlainomonas* could play a role in albedo-driven snow melt previously ascribed to *Chlamydomonas cf. nivalis* or *Sanguina*. *Chlainomonas* was not previously reported to be so prevalent in snow algae blooms of mixed composition. This could be because *Chlainomonas* is not distinguishable from *Chloromonas* with 18S primers in previous alpine HTS surveys [@Brown2016a,@Hamilton2017]. It is not reported from either pole, although this could be due to lack of sampling as well as lack of detect ability with HTS. -->

#

**Different alpine and subalpine snow algae communities**. *Sanguina* was more abundant in alpine sites, as well as one OTU of *Chlainomonas* and two OTUs of *Chloromonas*, while *Chloromonas krienitzii* and Chlainomonas OTU6 were found in high relative abundance below treeline (Fig. 2, 3). This consistent with the microscopy observations of @Hoham1979a and @Nedbalova2008a, who found primarily *Chlamydomonas cf. nivalis* above treeline and *Chloromonas* below treeline. However, in microscopy surveys we infrequently observed *Sanguina* below treeline, and *Chloromonas krienitzii* above treeline. We found higher diversity in samples from higher elevation, both at the ASV level and OTU level<!--Chao1 w fossil.-->, with several OTUs that were not found at lower elevations. *Raphidonema* and related ASVs were restricted to high alpine sites, with the highest relative abundance occurring in samples from August and September. This fits with the interpretation of @Stibal2005, who suggest that *Raphidonema* is an opportunistic soil algae that blows into snow fields from surrounding dirt to colonize the snow. However, rare taxa should not be discounted as playing an important role in ecosystem functioning [@Jousset2017].

Snowpack above and below treeline are very different habitats, with forest providing shade (and therefore less evaporation). The red pigment of *Sanguina* and *Chlainomonas* (astaxanthin) is likely an adaptation to high light [@Gorton2003;@Holzinger2006], and it has been suggested that *Sanguina* is adapted to the drier snowpacks that form at higher elevations [@Dial2018c].

#

**_Chloromonas krienitzii_ is green below, turns orange at surface**. On initial visits to Chloromonas krienitzii dominated sites, green snow was hidden 2 to 5 cm below the surface of the runnel (Supp. Fig. 4). Green snow was noticeably moister than surrounding snow, in some locations containing ice lenses not found in the surrounding snowpack. By June 6 the surface of the runnel had turned orange while the snow below remained green, which persisted on three subsequent visits until the snow melted. Cell morphology at the surface closely resembled *Chloromonas cf. brevispina* “Trochiscia” [@Matsuzaki2015] (Fig. 4<!-- which panel-->), while green subsurface snow was dominated by green biflagellate or bipolar quadriflagellate cells (Fig. 4). Both green and orange samples were dominated by reads assigned to *C. krienitzii*, with lower relative abundance of *Chlainomonas*, which was more abundant at the surface (Supp. Fig. 3). <!--We were surprised to find that --><!--H5 was dominated by *Chlainomonas* reads, but this sample contained few cells matching this morphology.-->

One possibility is that C. krienitzii cells orange pigmentation is triggered or an adaptation to higher light levels at the surface. Green cells below the surface would be protected from UV radiation<!--cite -->. Possibly, the green cells maintain an optimal light level through phototaxis, similar to marine phytoplankton<!--cite-->. Perhaps higher elevations typically have fewer low angle runnels that would allow a moist enough environment for *C. krienitzii*. It seems unlikely that nutrients from conifer needles would explain this, as several of our *C. krienitzii* dominated sites were in full sunlight, far from any trees.

#

**Snow algae blooms harbor novel diversity**. Several OTUs were not assigned to species level due to low similarity (Fig. 1). Two OTUs of Chloromonas, one OTU of Chlainomonas, and one OTU ("Algae OTU8") that had best BLAST matches to both Trebouxiophyceae and Chlorophyceae algae. Algae OTU8 has a best BLAST score of only 88%; this could represent a free-living algae found in snow, or possibly an endosymbiotic algae from a ciliate, as we noted many ciliates with green organelles in our microscopy samples. Sample X17 contained was dominated by spherical red cells superficially similar to *Sanguina*, this sample was dominated by reads of Chloromonas OTU7 and Algae OTU8. Cells in this sample had thicker cell wall, and some cells were slightly square-shaped (Fig. 4). 18S algae reads from this sample were equivocal, with best BLAST matches to both *Raphidonema* and *Chloromonas*. Currently there are just two species *Chlainomonas* in reference databases, but our results suggest there may be genetically distinct populations *Chlainomonas* at higher elevations (Fig. 1, Supp. Fig. 1, Supp. Fig 2). Although some of the high elevations ASVs were classified as *Chlainomonas rubra*, because we only included this species in our custom snow algae reference this could have biased our classifier, possibly causing over-classification [@Murali2018].

Although our rbcL primers successfully detected distinct OTUs, this is likely an underestimate of the true diversity. We did not detect any clustering within *Sanguina* using either 18S or rbcL, however when we included the homologous rbcL fragment of *S. aurentia* in our analysis it did not form a distinct OTU <!--this is where I am quesy about out ghetto OTU method-->. Furthermore, rbcL is highly differentiated in *Chloromonas* relative to other microalgae [@Nozaki2002], which could partly explain why *Sanguina* is less differentiated than *Chloromonas* or *Chlainomonas* (Supp. Fig. 1 and 2).

<!-- Runnels could provide habitat, e.g. liquid water for motility, nutrient mobilization, alternatively the algae could passively accumulate due to gravity. These green cells could swim up from the base of the snowpack, or be passively washed into the runnels from the surface. We observed some green snow ice lenses below these runnels, which could be due to higher snow moisture in the runnel, or due to the algae themselves through antifreeze production or albedo.  -->

<!-- Although all samples dominated by *C. krienitzii* were taken from elevations near or below treeline, our microscopy observations suggest that *C. krienitzii* blooms infrequently at higher elevations as well. Why *C. krienitzii* dominates below treeline is a mystery, since we observed it in high light clearings.  -->
<!-- Exposure to light can trigger astaxanthin production in Haematococcus pluvialis [@Shah2016], -->

# Discussion {-}

We found snow algae communities were typically dominated by either *Sanguina*, *Chlainomonas* or *Chloromonas*, with secondary abundance of *Chlainomonas* or *Chloromonas* in most samples. Most samples contained a sizable proportion of secondary or tertiary OTUs, although some some contained nearly 100% *Chlainomonas rubra*, *Sanguina nivaloides*, or *Chloromonas krienitzii*. The drivers of these differences between communities remain unclear; it is likely that within the melting snowpack there may be distinct microhabitats that impacts snow algae distribution, with varying water content with topography, different light levels in shade or at depth [@Gorton2007], different daily freeze-thaw cycles, or possibly varying nutrient levels with proximity to vegetation [@Hoham1976] or bedrock [@Kol1968a]. Further work needs to be done to uncover the micro-habitat niches within the snowpack that allow this differentiation, and what novel species the snow microbiome might contain.

<!-- Where to put these? -->
<!-- cross validation -->
<!-- Sample composition varied between our three data sets (rbcL, 18S, and cell count), although some robust correlations were found between them. *Sanguina* relative abundance was strongly correlated between data sets, between rbcL and 18s (Spearman's rho=0.78, p<0.001) and rbcL and "*Sanguina*-like" cell morphologies (Spearman's rho=0.92, p<0.001). *Chlainomonas* was strongly correlated with "*Chlainomonas*-like" cells (Spearman's rho=0.7, p<0.001), but IDTaxa did not identify any *Chlainomonas* in our 18S library (in spite of the fact that we included one sample in our 18S reference database). *Chloromonas krienitzii* was not correlated with cell count data (because green cells), nor any 18S ASVs or OTUs. Koliellaceae was not strongly correlated between 18S and rbcL datasets by either relative abundance data or presence absence. No strong correlations were found between Unk8 and taxa or OTUs in other datasets. Sample S4 was dominated by cell morphologies we classified as *Chloromonas cf. brevispina*, and this sample was dominated by reads best matching *Chloromonas krienitzii*. *Sanguina* and *Trebouxiophyceae* was in proportionally higher among 18S algae reads than in rbcL (respectively 40% vs 25%, and 0.9% vs 2%), while *Chloromonas* was lower in 18S than rbcL *Chloromonas* and *Chlainomonas* combined (24% vs 45%).  -->

<!-- mention our brevispina sample -->
<!-- Cell morphology also suggests rbcL underestimates diversity. Sample S4 was dominated by *C. krienitzii* reads, but the cell morphologies dominant in this samples were similar to *Chloromonas cf. brevispina* Cryocystis-type [@Matsuzaki2015], which is a distinct but closely related species to *C. krienitzii*. Consistent with @Nedbalova2008a, we found these cells in bronze snow in well-shaded forest sites. Likely our short-read rbcL fragment would not offer enough genetic resolution to differentiate between the two, with the genetic "signal" of *C. cf. brevispina* getting lost in the "noise" of haplotype variation in *C. krienitzii*. -->

<!-- primer bias -->
<!-- Primer bias likely impacted our estimates of relative abundance. Samples containing 80-90% *Sanguina* 18S were only 50-60% *Sanguina* rbcL. This is likely because we designed our rbcL primers off of *Chloromonas* and *Chlainomonas*, as *Sanguina* sequences were not available at the time, thus leading to their disproportionate representation. Perhaps the three uppermost outliers in Figure 2 (K7, K19.1, and G16.2) could be because these samples contained very little *Chloromonas* initially. Although *Chloromonas* and *Chlainomonas* may be artificially inflated by rbcL, they are likely a significant secondary component of *Sanguina* blooms, as our 18S data suggests, as well as microscopy. Other studies also found that *Chlamydomonas cf. nivalis* blooms contained lower abundance of *Chloromonas* [@Lutz2016a;@Terashima2017]. Furthermore extraction bias could likewise have resulted in over-representation of *Chloromonas* and *Chlainomonas*, since we observed that *Sanguina* were difficult to lyse by light microscopy cell counts, and contained lower DNA yields per dry weight than samples containing green cells. -->

<!-- Our rbcL amplicon library was dominated by reads assigned to *Chloromonas*, *Sanguina*, and *Chlainomonas* (Fig. 1). The most abundant species were *Chloromonas krienitzii*, *Chlainomonas rubra*, and *Sanguina nivaloides*. <!--exact percentage?--> <!--X% of reads were assigned to genus, and X% of reads were assigned to species. Many ASVs were not assigned to genus level by our taxonomic classifier IDTaxa, but were nevertheless closely related to *Chloromonas* (Fig. 1, Supp. Fig. 1 and 2). Although one OTU "Chloro7" was distant from other *Chloromonas* ASVs (Supp. Fig. 2), best BLAST matches were all *Chloromonas* snow algae (best match 92%). Similarly, Chloro9 had best BLAST matches to Chloromonas, with a best match of 94%. Another OTU "Unk8" was only assigned to domain level by IDTaxa, but top BLAST hits included algae from both Chlamydomonadaceae and Trebouxiophyceae, with a best match of 88% similarity. <!--This cluster did not correlate with any 18S OTUs. However, another unidentified clade (labeled as “Chlorophyceae X” in Figure 1) was highly correlated with an 18S OTU that was similarly placed taxonomically.-->

<!-- Did 18S identify any of the unknown clusters? What was in the unknown clusters? -->

<!-- Three green snow samples from above treeline were dominated by *Chloromonas* reads.  -->
<!-- Two Sanguina dominated samples contained high proportion of smaller orange cells... -->

<!-- move some of this to discussion? -->

 <!--However, the low abundance community was less differentiated between samples<!-- too discussion-y?-->  <!--How to mention the higher diversity at higher elevation? Chao with fossil pkg? There was greater genetic variation between high elevation samples than there was between lower elevation samples (Fig. 2b) this is because there were more samples? more sites? hard to make comparisons. <!-- ... becayse Chloromonas and Chlainomonas are closely related-->

<!-- should I go into Chlain genetics, above and below treeline? morph differences? If I get into the morphology, I am characterising new species. Maybe reviewers will want me to go back and flesh out the details of the unknown species, but for now I won't mention it to provoke their curiosity -->

<!-- todo, get 18S clade -> OTU that correlates to Chloro 7. get correl. -->

<!-- remind of the problem: we have all this diversity, where is it? what are the differences? -->

<!--Chloro3, what to make of this. Are we suggesting it is a single species? (NO)-->
<!-- why do we care if there are new species?-->

<!--minor point, whittle down-->

<!-- Red, orange, or green snow can be caused by different species of snow algae. Much of the snow algae literature has focussed on the eye-catching red snow prevalent throughout polar and alpine regions caused by red spherical cells [@Leya2013;@Segawa2018;@Lutz2016a]. Until recently, these cells were referred to in the literature as *Chlamydomonas cf. nivalis*, but were recently re-classified as *Sanguina nivaloides* [@Prochazkova2019]. Red snow can also be caused by *Chlainomonas*, whose distinctive large (40 $\mu$m diameter) quadriflagellate red cells can be found in waterlogged snow over alpine lakes [@Novis2008;@Prochazkova2018c]. Snow algae that lack secondary pigments can color snow green, which occurs in slushy wet snow in the Arctic and Antarctic [@Fogg1967;@Lutz2015;@Remias2013], or in subalpine forests [@Hoham1979b]. Green snow is now thought to be caused by different genera than in red snow, including *Microglena* [@Lutz2015], and various *Chloromonas* [Hoham1979b;@Remias2013a;@Terashima2017]. Two studies (in USA and Slovakia) found green and orange snow to be more common below treeline, while red snow is prevalent in the alpine [@Hoham1979a;@Nedbalova2008a]. <!-- add element of unknown to this, emphasize how little we know-->

<!-- Although  -->

<!-- Cell morphology can be ambiguous in snow algae taxonomy; different species can look nearly identical, and the same species can look completely different depending on the life stage. Recent molecular work shows that strains formerly lumped together as *Chloromonas. cf. nivalis* contain four genetically-distinct lineages [@Matsuzaki2019], while cells described as *C. cf. brevispina* contain at least two genetically distinct clades [@Matsuzaki2015].  -->

<!-- S.aurantia-->
<!-- sky27 and gar 2 couldbe had fewer Chloromonas, so there were no Chloromonas to preferentially amplify -->
<!-- haplotypes, a single ASV difference may not be biologically meaningful -->
<!--future directions: what is the unknowns? -->
<!-- Below treeline, trees shade snowpack and provide a potential source of nutrients through conifer needles [@Hoham1976]; furthermore the window of opportunity for snow algae growth is much shorter as snow melts out sooner at low elevations. -->

<!-- Problems with snow algae 18S amplicon high-throughput sequencing (HTS). Although these studies report taxonomy to the species level, outdated annotation in reference databases, algorithms prone to overclassification [@Murali2018], lack of cross-referenced marker genes and microscopy samples, primers targeting highly conserved regions, and arbitrary OTU thresholds [@Callahan2017] could have hidden or mislabelled taxa, potentially obscuring variation between samples.  -->
<!-- - Most species of *Chloromonas* are known from a single sample.  -->

<!-- blooms side by side different community, which sample? -->
<!-- rbcL underestimates non Chloromonas, less differentiated within Sanguina -->
 <!-- nes18.01, has chlainomonas, but visually more sanguina. primer bias, check 18s.  -->
 <!-- are the citrine cells sanguina? -->
 <!-- return to this in discussion, we found that green snow near red snow was Chloromonas, but we did see some green cells below the surface in transition to red, similar to reported in Prochazkova2019. Where these cells come from is a enduring mystery, how the bloom is seeded. We observed red snow in patches on the saddles, with no green below. The snow was shallow below the alpine green Chloromonas. -->

# Methods {-}

**Field sampling and microscopy.** We collected snow algae from mountains near Vancouver, British Columbia throughout the summer of 2018. To capture the extent of snow algae diversity in our local mountains we collected as many samples as possible from different elevations, dates, mountains, and micro-habitats within the snow <!--insert suppl table-->. In total we collected 310 colored snow samples from 13 different mountains on 33 different dates from elevations between 880 m and 2150 m above sea level. In early season we sampled at lower elevations, moving uphill as the snow algae bloom progressed upwards in elevation [@Takeuchi2013]. We observed the first small spots of red snow above treeline on June 20.

We scooped samples from visibly colored snow into 50 mL centrifuge tubes using sterile technique. To prevent melting en route to the lab we stored tubes in a bag of snow. Back in the lab, we melted each sample at room temperature on the bench and removed a 1 mL aliquot for light microscopy. Immediately after, samples were stored at -20 &deg;C for up to eight months until DNA extraction. Most samples were viewed using light microscopy, characterizing the dominant cell morphology in each sample. We prepared a slide of cells fixed in 2% gluteraldehyde, and quantified the most common cell morphologies with a cell count to 100, classifying each cell as either *Sanguina*, *Chloromonas cf. nivalis*, *Chloromonas cf. brevispina*, *Chloromonas krienitzii* *Chlainomonas*, or "other".

#

**DNA extraction and amplicon library preparation.** We selected 35 out of 310 samples for rbcL and 18s high-throughput amplicon sequencing. We chose these samples to represent the variation in date, elevation, geographic location, snow color, micro-habitat, and cell morphology. We freeze-dried these samples for up to 48 hours until until samples appeared completely desiccated. To allow water vapor to escape sample tubes while freeze drying we poked holes in each lid; as a safeguard against cross-contamination from airborne snow algae powder we wrapped tubes individually with paper towels. We mini-pestled 5 mg of sample at room temperature to physically rupture cell walls. Based on our observation that samples dominated by *Sanguina* cell morphologies had lower DNA yields, and had many intact cells following various lysis methods, we used 20 mg of these samples to increase our DNA yield.

To extract DNA from the crushed cells we added 800 mL 1x CTAB extraction buffer [@ColdSp2009], 1% B-mercaptoethanol, 5 $\mu$L each of proteinase K and RNAase A, and incubated these at 65 &deg;C for 30 minutes. We centrifuged samples at 10,000 g for 3 minutes to pellet cell debris, then added 700 $\mu$L of supernatant to an equal volume of 24:1 chloroform:isoamyl alcohol. We inverted samples to mix and centrifuged again at 12,000 g for 10 minutes. We precipitated DNA by transferring the top layer to 700 $\mu$L ice-cold ethanol, which we gently mixed, and spun over Qiagen DNA columns for 30 s at 15,000 g. We washed columns twice with 70% ethanol, and finally dissolved our DNA by spinning with 50 $\mu$L of sterile TE buffer. As a negative control we processed a sterile distilled water sample alongside each batch, treating it exactly the same as the other tubes.

We used rbcL as a marker gene, which offers high resolution between microalgae species [@Zou2016], and has reference data for snow algae available on GenBank. We also sequenced each sample with 18S primers, due to it's coverage across a wide range of taxa, and its ubiquity in reference databases. We designed rbcL primers to target an approximately 400 bp section of this gene, based on 20 snow algae rbcL GenBank sequences from *Chloromonas* and *Chlainomonas* (GenBank accession numbers AB434272.1, LC012752.1, LC012747.1, AF517072.1, LC012738.1, LC012739.1, AB434267.1, EU030690.1, LC360494.1, AJ001878.1, AB022225.1, DQ885964.2, DQ885962.1, AJ001879.1, AB022226.1, AB022530.1, LC012751.1, AB504764.1, EU030689.1, AB101508.1). *Sanguina* sequences were not included because they were not available at the time. We designed primers with PRIME(+) from the GCG package. These primers were rbcL369F (5’-GAA CGT GAC AAA TTA AAC AAA-3’) and rbcL870R (5’-ACC WGA YAD ACG WAG AGC TT-3’). To target 18S we used Euk1181 (5'-TTA ATT TGA CTC AAC RCG GG-3') and Euk1624 (5'-CGG GCG GTG TGT ACA AAG G-3') [@Wang2014].

We constructed our amplicon library using a two-step PCR [@Meyer2010]. In the first PCR we amplified template DNA using our primers attached to a universal adapter, and in the second PCR we re-amplified that product to attach a 6 bp index to the universal adapter at the 3’ end. The first PCR total volume was 25 $\mu$L, consisting of 1 $\mu$L template, 12.5 $\mu$L Q5 high-fidelity 2X MM (New England BioLabs), 1.25 $\mu$L each of forward and reverse primer, and 9 $\mu$L of ddH2O. The second PCR was the same except we reduced our reaction volume to 20 $\mu$L by using only 5 $\mu$L of ddH2O. The cycling conditions were the same for both primer pairs for the first PCR, with an initial denaturation at 98 &deg;C for 30 s, followed by 30 cycles of 98 &deg;C for 5 s, 58 &deg;C for 10 s, and 72 &deg;C for 25 s, with a final extension at 72 &deg;C for 2 minutes. For the second indexing PCR we started with an initial denaturation at 98 &deg;C for 30 s, then 10 cycles of 98 &deg;C for 10 s, 65 &deg;C for 30 s, and 72 &deg;C for 30 s with a final denaturation of 72 &deg;C for 5 min. After each PCR, we purified using Agencourt AMPure XP kit (Beckman Coulter). We quantified final DNA concentration with Qubit (Thermo Fisher), and standardized sample concentration for pooling. The pooled library was then loaded and run on an Illumina MiSeq V3 kit.

#

**Bioinformatic processing.** Reads were demultiplexing using cutadapt [@Martin2011]. Samples were filtered and trimmed, errors removed, dereplicated, pair end reads merged, and chimeras removed in DADA2 [@Callahan2017]. <!-- singletons, failed samples removed, remove sequences based on negative controls, adjust for 0's, -->

We assigned taxonomy using a custom reference databases for both rbcL and 18S, since snow algae are not well-represented in SILVA or PR2, and there is no curated database for rbcL. We made our rbcL database by downloading 30,865 algae rbcL sequences from GenBank with the query “rbcl[gene] AND (algae OR chlorophyta OR trebouxiophyceae) NOT (18s OR ribosomal OR psaB OR atpB) AND 300:2000[slen] NOT plasmid NOT unverified NOT mRNA NOT bacteria NOT mitochondrion”. We removed ambiguous snow algae annotations based on the most recent snow algae phylogeny [@Matsuzaki2019]. Because entries labelled as *Chloromonas cf. brevispina* and *Chloromonas cf. nivalis* are polyphyletic we re-labeled these as “unassigned *Chloromonas*”. <!--Where multiple sequences were present for a single species we generated a consensus sequence to avoid imbalance where some groups have far more representatives than others [@Murali2018] --><!--To  re-labeled GenBank accessions LC012733.1 and LC012735.1 as Chloromonas krienitzii, LC012734.1 as Chloromonas “Cryocystis”-type, and LC012744.1 to Chloromonas miwae, and LC012743 and LC012745 to “Chloromonas Group_B” 6`.--> Species designated by morphology alone that were not genetically different from other species in their genera were annotated only to genus level. <!-- all Chlainomonas to genus level only --> We assigned rbcL ASV taxonomy using IDTaxa with threshold = 50.

To assign 18S reads we initially ran our ASVs against SILVA [@Quast2013], and selected ASVs assigned to Chlorophyta. We then ran this subset of ASVs on our custom 18S snow algae database based on the following GenBank query: "(18S OR ribosomal)[gene] AND (chloromonas OR chlainomonas OR sanguina OR raphidonema OR KMY-2018) NOT (rbcL OR psaB OR atpB) AND 300:3500[slen] NOT plasmid NOT unverified NOT mRNA NOT bacteria NOT mitochondrion". This database was edited to reflect the most recent snow algae taxonomy and remove ambiguous annotations [@Matsuzaki2019;@Prochazkova2018c]. We assigned taxonomy using IDTaxa with threshold = 50.

We assigned OTUs with t-SNE, based on visually distinct clusters [@Graham-Rowe1999] <!--BibTeX header messed up-->(Fig. 1). We split the largest cluster (containing the vast majority of the *Chloromonas* reads) into three clusters, one containing all ASVs assigned to *Chloromonas krienitzii*, and one containing all ASVs assigned to *Chloromonas muramotoi* and *Chloromonas hohamii*. To highlight our observation that one clade of *Chlainomonas* were only found above treeline (Suppl. Fig. 1), we split *Chlainomonas* into two OTUs. The OTUs produced by this naive method were roughly consistent across different methods of visualization (Suppl. Fig. 1 and 2). <!--Other OTU classifiers such swarm [@Mahe2014] are optimized for 18S data, and we found swarm did a poor job of clustering when visualised with MDS, t-SNE, and phylogenetic trees, with various values of d. <!--and CROP [@Chen2013].-->
<!--OTU clustering with t-SNE: based on poor results with other classifiers, we used tsne. since tsne is known to find patterns where there are none depending on the perplexity, we ran this test with perplexity ranging from 5-100, and found the same large scale clusters each time. -->

**Software information.** All analysis was conducted in R 3.6.1 with tidyverse 1.2.1, with dada2 1.12.1, DECIPHER 2.12.0, Rtsne 0.15, GUniFrac 1.1, and vegan 2.5-6. Phylogenetic analysis was carried out in IQTree [@Nguyen2015] and visualized with ggtree [@Yu2017].

**Data availability.** All raw fastq files are freely available on the European Nucleotide Archive under the project accession PRJEB34539. All scripts are available at cengstro.github.io/projects/sa_rbcl.

##### pagebreak

# Acknowledgments {-}

We wish to thank Leah Tooman (Simon Fraser University) for assistance with sequencing, and Chris Rushton (Simon Fraser University) for assistance with bioinformatics. <!--We also wish to thank Forest Rohwer (UCSD) for the inspiration that led to this project.--> This project was funded with a Sector Innovation Grant from Genome BC (SIP016), and a NSERC Individual Discovery Grant, both awarded to LQ.

# Author Contributions {-}

<!--

When determining authorship the following criteria should be observed:

 - Substantial contributions to the conception or design of the work; or the
   acquisition, analysis, or interpretation of data for the work; AND
 - Drafting the work or revising it critically for important intellectual
   content; AND
 - Final approval of the version to be published ; AND
 - Agreement to be accountable for all aspects of the work in ensuring that
   questions related to the accuracy or integrity of any part of the work are
   appropriately investigated and resolved.

Contributors who meet fewer than all 4 of the above criteria for authorship
should not be listed as authors, but they should be acknowledged.
(http://www.icmje.org/roles_a.html)

-->

CE, LQ, and KY designed this study. Samples collected by CE with assistance from KY and LQ. KY and CE prepped samples for sequencing. All analyses were completed by CE and KY. All authors discussed the results and contributed to the final manuscript. CE wrote the manuscript with major input from LQ and KY.
<!-- The statement about the authors and contributors can be up to several sentences -->
<!-- long, describing the tasks of individual authors referred to by their initials -->
<!-- and should be included at the end of the manuscript before the References -->
<!-- section. -->

# Additional information {-}

### *Supplementary information*.

### *Competing financial interests.* The authors declare that the research was conducted in the absence of any
commercial or financial relationships that could be construed as a potential
conflict of interest.

##### pagebreak

# References {-}
<div id="refs"></div>

# Figures {-}

<br>
<br>

![Figure 1. t-SNE clustering of snow algae rbcL amplicon sequence variants (ASVs). Colored points correspond to genus and species level assignment, grey points were not assigned at or below genus level. Labelled ellipses show OTUs. Perplexity = 30, theta = 0.5](/home/cengstro/projects/sa_rbcl/figs/asv_tsne_fig.png)

<!-- todo: italics, underscores, ChlorKri5, ChlorMura5, put ChloroKri4 on top of Chloro3, so its obvious that these aren't part of both OTUs.-->

##### pagebreak

![Figure 2. Sample ordination plots. A. NMDS comparing sample similarity using OTUs, shaded by elevation. Vectors indicate relative contribution of taxa driving differences between samples. Stress = 0.183. B. NMDS showing weighted UniFrac distances between snow algae samples. Stress = 0.11.](/home/cengstro/projects/sa_rbcl/figs/sample_ordination.png){width=13.01cm, height=18.56cm}

<!-- add dom morph groupings for UniFrac to connect with heatmap.  adjust size.  -->
<!-- change axis labs of b, MDS to NMDS-->

##### pagebreak

![Figure 3. Relative abundance heatmap of snow algae OTUs in each sample. Weight of color corresponds to the relative abundance of that OTU within a sample. Samples (rows) are grouped by the dominant cell morphology within that sample as determined by light microscopy. Within each morphology group samples are ordered by elevation, from low on the bottom to high on top. Sample IDs follow the convention: letters represent the mountain, and the number represents the order in which it was collected. Samples collected on the same day are labelled as .1, .2, and so on.](/home/cengstro/projects/sa_rbcl/figs/heatmap.png){width=12.88cm, height=17.17cm}

<!-- add elev range, date range. dom cell morph text to grey. lump chlainomonas categories (let it be a surprise)-->
<!-- add treeline -->
```{r, echo = F, include=F}
# otu_heat2 %>% 
#   group_by(bloom_type) %>% 
#   summarise(min(elev_m), max(elev_m))
# otu_heat2 %>% 
#   group_by(bloom_type) %>% 
#   summarise(min(date), max(date))
```

##### pagebreak

![Figure 4. Select microphotographs of snow algae. All photos are taken with light microscopy, except (f) which was taken with DIC. Scale bars represent 30 $\mu$m. (a) H5 (b) S10.2. (c) S4. (d) K2. (e) S10.1. (f) G16.1. (g) N12.1. (h) P20.1.](/home/cengstro/projects/sa_rbcl/figs/fig4.png){width=13.51cm, height=17.49cm}
<!-- order by elev, add sax1 -->

<!-- ![Figure 5. Relative abundance barplots comparing community composition in surface orange snow and green snow below in three samples.] -->
<!-- choose photo, microphotos, rethink colors (make consistant w heatmap) -->

##### pagebreak

# Supplementary Data {-}

![Supplementary Figure 1. Maximum likelihood tree (IQTree) of snow algae rbcL ASVs. Colors correspond to OTU. Edges with bootstrap values < 80 are displayed as dotted lines.](/home/cengstro/projects/sa_rbcl/figs/ggtree.png)

##### pagebreak

![Supplementary Figure 2. MDS plot of rbcL ASVs, colored by taxa. Ellipses contain all ASVs within an OTU. Stress =  ](/home/cengstro/projects/sa_rbcl/figs/asv_mds.png)

<!-- redo with NMDS, include stress vals -->
##### pagebreak

![Supplementary Figure 3. Relative abundance barplots comparing community composition in surface orange snow and subsurface green snow. * designates depth and surface samples collected on different days.](/home/cengstro/projects/sa_rbcl/figs/draft1/depth.png)

##### pagebreak

<!-- ![Supplementary Figure 4. Photos showing location of *Chloromonas krienitzii* dominated samples.]() -->

<!-- word to template: *[:alpha:] to \*, and *[:space:] to } -->