### 2019-10-24 code/community_comparison.Rmd
transfer files over from old rproj

---
title: "Community comparison"
output: html_notebook
---

```{r libs, echo=F}
library(tidyverse)
library(GUniFrac)
library(vegan)
library(ggthemes)
```

```{r set}
theme_set(theme_few())
set.seed(123)
```

```{r input, echo=F}
rbcl_rel_abund <- read_csv(here("data","clean","rbcl_rel_abund_clean.csv"))
idtax <- read_csv(here("data","clean","rbcl_taxonomy_clean.csv"))
field <- read_csv(here("data","clean","field_clean.csv"))

rbcl_asv_id_key <- read_csv(here("output","keys","rbcl_asv_key.csv"))
rbcl_asv_otu_key <- read_csv(here("output","keys","asv_otu_key.csv"))
sample_aliases <- read_csv(here("data","sample_aliases.csv"))
rbcl.tree <- read.tree(here("output","tree","rbcl","asv613.fasta.treefile"))
```

Unifrac rbcL
```{r unifrac}
rbcl.asv.rel.abund.wide <- rbcl_rel_abund %>% 
  select(asv_id, sample_id, rel_abund) %>% 
  pivot_wider(names_from = asv_id, values_from = rel_abund, values_fill = list(rel_abund = 0)) %>% 
  column_to_rownames("sample_id")

# rooted tree required, root by one of my Raphidonema asvs
rbcl.tree <- rbcl.tree %>% root(outgroup = "559", resolve.root=T)

unifracs <- rbcl.asv.rel.abund.wide %>% 
  GUniFrac::GUniFrac(rbcl.tree, alpha=c(0, 0.5, 1))
unifrac <- unifracs$unifracs[, , "d_1"]

# plot results as NMDS
nmds <- metaMDS(unifrac, autotransform = F)
nmds # stress
unifrac_plot_data <- nmds %>% 
  scores() %>% 
  as_tibble(rownames = "sample_id") %>% 
  left_join(field, by="sample_id")

# plot unifrac, colored by elevation
uf_plot <- ggplot(unifrac_plot_data, aes(x=NMDS2, y=NMDS1)) +
  geom_point(aes(color=elev_m), size=4) +
  scale_color_gradientn(colours = terrain.colors(8)[1:7]) +
  labs(color="Elevation (m)") +
  geom_text(data = unifrac_plot_data %>%   # add sample ID labels
              left_join(sample_aliases, by="sample_id"),
            aes(label = alias), size=3)  ######### change to alias for final version. edit overlap txt in svg
uf_plot
```

# rbcL relative abundance plot
```{r barplot}
rbcl_stack_barplot_tbl <- rbcl_rel_abund %>% 
  left_join(rbcl_asv_otu_key, by="asv_id") %>% 
  left_join(field, by="sample_id") %>% 
  left_join(sample_aliases, by="sample_id") %>% 
  select(alias, rel_abund, otu_name, elev_m) %>% 
  group_by(alias, otu_name, elev_m) %>% 
  summarise(rel_abund = sum(rel_abund)) %>% 
  ungroup()

rbcl_stack_barplot_tbl %>% distinct(otu_name)

# sanity check samples rel abund sum to 1
rbcl_stack_barplot_tbl %>%
  group_by(alias) %>% 
  summarise(sum(rel_abund))

# order OTUs
# otu.order <- c("Chloromonas", "Sanguina A","Trebouxiophyceae J")

stacked_barplot <- ggplot(rbcl_stack_barplot_tbl, aes(x = alias %>% fct_reorder(elev_m), y = rel_abund, fill=otu_name)) +
  geom_bar(position="stack",stat="identity") +
  coord_flip() +
  ylab("Sample ID") +
  xlab("Relative abundance") +
  labs(fill = "OTU")
stacked_barplot

```

```{r output}
ggsave(plot = uf_plot, filename = here("figs","rbcl_unifrac_nmds.pdf"), height = 6, width = 8)
ggsave(plot = stacked_barplot, filename = here("figs","rbcl_otu_rel_abund.pdf"), height = 6, width = 8)
```