### 2019-10-29 code/wrangle/wrangle_18s_rel_abund.Rmd

---
title: "Wrangle 18s rel abund data"
output: html_notebook
---

```{r libs, results="hide"}
library(tidyverse)
library(here)
library(metagenomeSeq)
```

Read in file, and view contents

```{r input, results="hide"}
path_18s_ra <- here("data/raw/ky_18s_seqtab.nochim_April2019.txt")
raw_18s_ra <- read.csv(rel_abund_18S_path, sep=" ", stringsAsFactors = F)

# also read in the rbcl rel abund tidy data, since I'm only using the 18s data as a cross reference I can filter by this
path_rbcl_ra <- here("data/clean/rbcl_rel_abund_clean.csv")
clean_rbcl_ra <- read_csv(path_rbcl_ra)
```

Object naming convention: 
- first what the objects is in its local context, then any globel identifiers
- I'll refer to 18S relative abundance table object as "18s_ra"

Rename sample IDs
- remove MiSeq lane identifier ("S" followed by a number)
```{r}
head(raw_18s_ra)
renamed_sids_18s_ra <- raw_18s_ra %>%
  as_tibble(rownames = "sample_id") %>% 
  mutate(sample_id = sample_id %>% 
           str_remove("_S.*_") %>% # remove _S*_ ending
           str_replace_all("-",".")) #replace hyphen with period
head(renamed_sids_18s_ra)
```

Transpose so each ASV a row, each sample a column
```{r}
transposed_18s_ra <- renamed_sids_18s_ra %>% 
  column_to_rownames("sample_id") %>% 
  t() %>% 
  as_tibble(rownames = "asv_sequence")
head(transposed_18s_ra)
```

Remove actual sequences from tbl, set aside in a seperate tbl

First check each ASV unique
```{r asv_seq_key}
selected_asvs <- transposed_18s_ra %>% 
  select(asv_sequence)
distinct_asvs <- transposed_18s_ra %>% 
  distinct(asv_sequence)
identical(selected_asvs, distinct_asvs)
```

Make a key table
```{r}
asv_seq_key_18s <- distinct_asvs %>% 
  rownames_to_column("asv_id") %>% 
  mutate(asv_id = asv_id %>% 
           as.character() %>% 
           str_pad(width=3, pad="0"))
glimpse(asv_seq_key_18s)
```

Remove the sequence from the main table
```{r}
seq_removed_18s_ra <- asv_seq_key_18s %>% 
  left_join(transposed_18s_ra, by="asv_sequence") %>% # left_join ensures that we are matching id with sequence by the master key
  select(-asv_sequence) # remove sequence, put asv_id first

head(seq_removed_18s_ra)
```

Normalize

Do this before we start to filter and select names

Convert to metagemoneSeq format
```{r normalize}
# mr_obj_18s_ra <- seq_removed_18s_ra %>% 
#   column_to_rownames("asv_id") %>% 
#   newMRexperiment()
# 
# mr_obj_18s_ra
```

```{r normalize2}
# 
# # calculate percentile by which to normalize
# p <- cumNormStatFast(mr_obj_18s_ra)
# 
# 
# # normed_18s_ra <- 
#   
# mr_obj_18s_ra %>% 
#   
#   
#   cumNorm(
#   MRcounts(norm=T) %>% 
#   as_tibble(rownames = "asv_id")
# 
# %>% 
#   gather("sample_id","norm_reads", -asv_id)
# 
# rbcl_norm <- rbcl_norm %>% 
#   filter(norm_reads!=0) %>% 
#   group_by(sample_id) %>% 
#   mutate(norm_rel_abund = norm_reads/sum(norm_reads))
# 
# rbcl_rel_abund %>% 
#   left_join(rbcl_norm, by=c("sample_id","asv_id")) %>% 
#   select(asv_id, sample_id, rel_abund, norm_rel_abund) %>% 
#   mutate(same = if_else(rel_abund == norm_rel_abund, T, F),
#          diff = rel_abund - norm_rel_abund) %>% 
#   arrange(-diff)
# filter(same==F)
```

Convert to long format
```{r long}
long_18s_ra <- seq_removed_18s_ra %>% 
  pivot_longer(cols = -asv_id, names_to = "sample_id", values_to = "n_reads") %>% 
  filter(n_reads != 0) # now that its in long format we can remove 0s

glimpse(long_18s_ra)

```

Calculate relative abundance as percentage
```{r}
as_percent_18s_ra <- long_18s_ra %>% 
  group_by(sample_id) %>% 
  mutate(rel_abund = n_reads/sum(n_reads)) %>% 
  ungroup()
glimpse(as_percent_18s_ra)
```

# Set aside negative controls

look for duplicate sample ids
```{r}
as_percent_18s_ra %>% 
  distinct(sample_id)
```

set aside duplicate sample ids
```{r}
duplicate_sample_18s_ra <- as_percent_18s_ra %>%
  filter(sample_id %>% str_detect("gar18.01|sax18.01|sey18.25|sey18.74|sky18.12"))
glimpse(duplicate_sample_18s_ra)
```

remove duplicates from my main analysis tbl
```{r}
no_dups_18s_ra <- as_percent_18s_ra %>% 
  filter(!(sample_id %in% c("gar18.01.p2", "sax18.01.p1", "sey18.25.p1", "sey18.74.p1", "sky18.12.p1")))
glimpse(no_dups_18s_ra)
```

remove the "p" from the sample id
```{r}
sample_id_fixed_18s_ra <- no_dups_18s_ra %>% 
  mutate(sample_id = str_remove(sample_id, "\\.p[:digit:]"))
glimpse(sample_id_fixed_18s_ra)
```

remove negative controls
```{r}
clean_18s_ra <- sample_id_fixed_18s_ra %>% 
  filter(sample_id %in% clean_rbcl_ra$sample_id)

glimpse(clean_18s_ra)
```

Check number of reads per sample
```{r}
clean_18s_ra %>% 
  group_by(sample_id) %>% 
  summarise(sum_n = sum(n_reads)) %>% 
  arrange(sum_n)
```

write clean csv as output
```{r}
clean_18s_ra_path <- here("data/clean/clean_18s_ra.csv")
write_csv(clean_18s_ra, clean_18s_ra_path)
```