### 2019-10-29 code/wrangle/wrangle_cellct.Rmd
initial commit

---
title: "Clean cell count data"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)
```

```{r}
path_cellct <- here("data/raw/cellct_2018.csv")
raw_cellct <- read_csv(path_cellct)
glimpse(raw_cellct)
```

```{r}
thresh <- 50 # set this much lower than 100, to keep some counts with high fungi relative abundance

cellct <- raw_cellct %>%
  # drop non-algae categories
  select(-three_arm:-biofilm) %>% 
  pivot_longer(cols = -sample_id, names_to="morpho_sp", values_to="count") %>%
  filter(count != 0) %>% 
  drop_na() %>% 
  mutate(morpho_sp = case_when( str_detect(morpho_sp, "green")~"Green cell", # lump all green categories
                                morpho_sp=="orb"~"Sanguina nivaloides", 
                                morpho_sp=="ruby"~"Sanguina nivaloides", # assuming Sanguina grows turrets 
                                morpho_sp=="balloon"~"Chlainomonas",
                                morpho_sp=="tangerine"~"Chloromonas krienitzii",
                                morpho_sp=="hedgehog"~"Chloromonas cf. brevispina",
                                morpho_sp=="lemon"~"Chloromonas cf. nivalis",
                                morpho_sp=="spiny"~"Other",
                                morpho_sp=="oval"~"Other",
                                morpho_sp=="zuke"~"Other",
                                morpho_sp=="halo"~"Other",
                                morpho_sp=="citrine"~"Sanguina nivaloides",
                                TRUE ~ morpho_sp)) %>% 
  # drop samples with count < thresh
  group_by(sample_id) %>% 
  mutate(count_total = sum(count)) %>% 
  filter(count_total > thresh) %>% 
  # calculate percentage
  mutate(percent_of_sample = count / count_total) %>% 
  ungroup()

head(cellct)
```

```{r}
path_to_write <- here("data/clean/clean.cellct.csv")
write_csv(cellct, path_to_write)
```