### 2019-10-30 code/make_18s_algae_fasta.Rmd
init

---
title: "Select Algal 18S ASVs "
output: html_notebook
---

```{r}
library(tidyverse)
library(DECIPHER)
```

```{r}
key_path <- here("intermediate_output/asv_seq_key_18s.csv")
key <- read_csv(key_path)

path <- here("data/raw/18s_assignments_silva.csv")
assignments <- read_csv(path)
assignments
```

# Only Algae

Select Chlorophyta ASVs (according to SILVA)
```{r}
algae_18s_seqs <- assignments %>% 
  dplyr::rename(asv_sequence = asv) %>% 
  filter(V3 %>% str_detect("Chlorophyta")) %>% 
  # get the asv_id, for labelling the tree tips
  left_join(key, by="asv_sequence") %>% 
  select(asv_id, asv_sequence) %>% 
  deframe() %>% # to vector
  DNAStringSet() 
algae_18s_seqs
```
Align
```{r}
aligns <- algae_18s_seqs %>%
  AlignSeqs(verbose=F)

BrowseSeqs(aligns, highlight = T)
```

Remove ASVs with gaps
```{r}
gappy_asvs <- c("508", "570", "581", "586", "587", "591")

quality_aligns <- aligns[!(names(aligns) %in% gappy_asvs)]

# check
length(quality_aligns) == length(aligns) - length(gappy_asvs)
```

# try with all ASVs

Make another fasta with all 18S ASVs,
```{r}
seqs <- assignments %>% 
  dplyr::rename(asv_sequence = asv) %>% 
  # get the asv_id, for labelling the tree tips
  left_join(key, by="asv_sequence") %>% 
  select(asv_id, asv_sequence) %>% 
  deframe() %>% # to vector
  DNAStringSet()
seqs
```

Remove gappy seqs
```{r}
seqs <- seqs[!(names(seqs) %in% gappy_asvs)]
seqs
```

Align
```{r}
all_aligns <- seqs %>%
  AlignSeqs(verbose=F)

BrowseSeqs(all_aligns, highlight = T)
```

Not the best looking alignment I've ever seen, but roll with it

Write fastas
```{r}
path_write <- here("intermediate_output/algae_18s_seqs.fasta")
writeXStringSet(quality_aligns, path_write)

path_all <- here("intermediate_output/all_18s_aligns.fasta")
writeXStringSet(all_aligns, path_all)
```