### 2019-10-31 code/depth_comparison.Rmd
split up wrangle tasks into scripts, made seperate csvs

---
title: "Depth sample comparison"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)
```

```{r include=F}
rbcl_rel_abund <- read_csv(here("data/"))
s18_all_rel_abund

rbcl_tax
s18_tax
```