### 2019-10-31 code/wrangle/wrangle_18s.Rmd

---
title: "R Notebook"
output: html_notebook
---

# asdf
test

```{r libs}
library(tidyverse)
library(here)
library(metagenomeSeq)
print("a")
```

Read in file, and view contents

Object naming convention: first what the objects is in its local context, then any globel identifiers

I'll refer to 18S relative abundance table object as "18s_ra"

```{r input}
path_18s_ra <- here("data/raw/ky_18s_seqtab.nochim_April2019.txt")
raw_18s_ra <- read.csv(rel_abund_18S_path, sep=" ", stringsAsFactors = F)
head(raw_18s_ra)
```

Transpose so each ASV a row, each sample a column
```{r}
transposed_18s_ra <- raw_18s_ra %>%
  t() %>% 
  as_tibble(rownames = "asv_sequence")
head(transposed_18s_ra)
```

Remove actual sequences from tbl, set aside in a seperate tbl

First check each ASV unique
```{r asv_seq_key}
selected_asvs <- transposed_18s_ra %>% 
  select(asv_sequence)
distinct_asvs <- transposed_18s_ra %>% 
  distinct(asv_sequence)
identical(selected_asvs, distinct_asvs)
```

Make a key table
```{r}
asv_seq_key_18s <- distinct_asvs %>% 
  rownames_to_column("asv_id") %>% 
  mutate(asv_id = asv_id %>% 
           as.character() %>% 
           str_pad(width=3, pad="0"))
nrow(asv_seq_key_18s)
```

Remove the sequence from the main table
```{r}
seq_removed_18s_ra <- asv_seq_key_18s %>% 
  left_join(transposed_18s_ra, by="asv_sequence") %>% # left_join ensures that we are matching id with sequence by the master key
  select(-asv_sequence) # remove sequence, put asv_id first

head(seq_removed_18s_ra)
```

Normalize

Do this before we start to filter and select names

Convert to metagemoneSeq format
```{r normalize}
# mr_obj_18s_ra <- seq_removed_18s_ra %>% 
#   column_to_rownames("asv_id") %>% 
#   newMRexperiment()
# 
# mr_obj_18s_ra
```

```{r normalize2}
# 
# # calculate percentile by which to normalize
# p <- cumNormStatFast(mr_obj_18s_ra)
# 
# 
# # normed_18s_ra <- 
#   
# mr_obj_18s_ra %>% 
#   
#   
#   cumNorm(
#   MRcounts(norm=T) %>% 
#   as_tibble(rownames = "asv_id")
# 
# %>% 
#   gather("sample_id","norm_reads", -asv_id)
# 
# rbcl_norm <- rbcl_norm %>% 
#   filter(norm_reads!=0) %>% 
#   group_by(sample_id) %>% 
#   mutate(norm_rel_abund = norm_reads/sum(norm_reads))
# 
# rbcl_rel_abund %>% 
#   left_join(rbcl_norm, by=c("sample_id","asv_id")) %>% 
#   select(asv_id, sample_id, rel_abund, norm_rel_abund) %>% 
#   mutate(same = if_else(rel_abund == norm_rel_abund, T, F),
#          diff = rel_abund - norm_rel_abund) %>% 
#   arrange(-diff)
# filter(same==F)
```

Convert to long format
```{r long}
long_18s_ra <- seq_removed_18s_ra %>% 
  pivot_longer(cols = -asv_id, names_to = "sample_id", values_to = "n_reads")

head(long_18s_ra)
nrow(long_18s_ra)
```

Format sample IDs properly (thanks to our poor organization during sequencing)
```{r reformat_sample_id}
fixed_sample_ids_18s_ra <- long_18s_ra %>% 
  mutate(sample_id = 
           # remove underscores
           str_extract(sample_id, ".*(?=_S.*_)") %>% 
           # replace hyphen
           str_replace_all("-","."), 
         # parse protocol version used from the sample ID col
         protocol = str_extract(sample_id, "(?<=[:alpha:]{3}18\\..{2}\\.p)[:digit:]"), 
         # remove protocol identifier from the sample ID col
         sample_id = str_replace(sample_id, ".p[:digit:]",""), 
         # protocol 3 was default, so can assume that these remaining are p3
         protocol = replace_na(protocol, 3))
fixed_sample_ids_18s_ra %>% 
  select(sample_id) %>% 
  distinct()
```

Calculate relative abundance as percentage
```{r}
as_percent_18s_ra <- fixed_sample_ids_18s_ra %>% 
  group_by(sample_id, protocol) %>% 
  mutate(rel_abund = n_reads/sum(n_reads)) %>% 
  ungroup()
as_percent_18s_ra
```

Set aside negative controls, and