### 2019-10-31 code/wrangle/old/wrangle.Rmd
moved these into "old" in case I need anything

---
title: "Wrangle"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)
```
Read in DADA2 output files from server, and sample site data
```{r input}
# rbcl asv table, and taxonomy table
seq.rbcl.2019 <- read.csv(here("data","raw","ky_rbcL_seqtab.nochim.txt"), sep=" ", stringsAsFactors = F) ##### KY saved as txt %^&*
idtax <- read_csv(here("data","raw","assignments_9-18.csv")) # thresh = 50

# 18s
asv.tbl.18s <- read.csv(here("data","raw","ky_18s_seqtab.nochim_April2019.txt"), sep=" ", stringsAsFactors = F)
s18_taxa_silva <- read_csv(here("data","raw","18s_assignments_silva.csv")) # thresh 40 ########### redo with thresh = 50 for consistancy

# site metadata
field <- read_csv(here("data","raw","field_2018.csv"))
```

Wrangle rbcL rel abund
```{r rbcl_rel_abund}
# rbcL relative abund --------------------

unfilt_rbcl_rel_abund <- seq.rbcl.2019 %>%
  t() %>% 
  as_tibble(rownames = "asv") %>% 
  gather(key = "sample_id", val = "n_reads", -asv)

# format sample ids properly
unfilt_rbcl_rel_abund <- unfilt_rbcl_rel_abund %>% 
  mutate(sample_id = str_extract(sample_id, ".*(?=_S.*_)") %>% # remove underscores
           str_replace_all("-","."), # replace hyphen
         protocol = str_extract(sample_id, "(?<=[:alpha:]{3}18\\..{2}\\.p)[:digit:]"), # parse protocol version used
         sample_id = str_replace(sample_id, ".p[:digit:]",""), # remove p1
         protocol = replace_na(protocol, 3)) # replace NA with 3

# compute new variable "rel_abund" 
############### consider normalizing with metagenomeSeq ( i think this weights for 0s somehow...)

unfilt_rbcl_rel_abund <- unfilt_rbcl_rel_abund %>% 
  group_by(sample_id, protocol) %>% 
  mutate(rel_abund = n_reads/sum(n_reads)) %>% 
  ungroup()
unfilt_rbcl_rel_abund
```

Seperate out controls, and samples with low reads
```{r controls}
# set aside and remove protocol controls
protocol.ids <- unfilt_rbcl_rel_abund %>% 
  select(sample_id, protocol) %>% 
  distinct() %>% 
  dplyr::count(sample_id) %>% 
  filter(n>1) %>% # choose if multiple protocols associated with a sample
  pull(sample_id)
protocol_comparison <- unfilt_rbcl_rel_abund %>% 
  filter(sample_id %in% protocol.ids)

# manually look for "protocol" samples to remove
protocol_comparison %>% 
  select(sample_id, protocol) %>% 
  distinct() %>% 
  group_by(sample_id) %>% 
  mutate(min=as.numeric(protocol) %>% min() %>% as.character(), 
         is_min = if_else(protocol == min, TRUE, FALSE)) %>% 
  filter(is_min == TRUE) %>% 
  select(1,2)
# ... and remove them from the main tbl manually
rbcl_rel_abund <- unfilt_rbcl_rel_abund %>%
  filter(!(sample_id == "gar18.01" & protocol == "2"),
         !(sample_id == "sax18.01" & protocol == "1"),
         !(sample_id == "sey18.25" & protocol == "1"),
         !(sample_id == "sey18.74" & protocol == "1"),
         !(sample_id == "sky18.12" & protocol == "1") )

# water negative controls
water_ctrls <- rbcl_rel_abund %>% 
  filter(sample_id %>% str_detect("water"))
rbcl_rel_abund <- rbcl_rel_abund %>% 
  filter(!(sample_id %>% str_detect("water")))

# white snow samples
white_snow_samples <- rbcl_rel_abund %>% 
  filter(sample_id %in% c("bre18.17", "bre18.22.bb"))
rbcl_rel_abund <- rbcl_rel_abund %>% 
  filter( !(sample_id %in% c("bre18.17", "bre18.22.bb")))

# scraped vs unscraped comparison
scrape_comparison <- rbcl_rel_abund %>% 
  filter(sample_id %>% str_detect("sky18.24"))
rbcl_rel_abund <- rbcl_rel_abund %>% 
  filter(!(sample_id == "sky18.24sc"))

# set aside depth samples
depth_sample_rel_abund <- rbcl_rel_abund %>% 
  filter(sample_id %in% c("hol18.42", "hol18.42b", "sey18.66", "sey18.66b", "hol18.21", "hol18.39d"))
rbcl_rel_abund <- rbcl_rel_abund %>% 
  filter(!(sample_id %in% c("hol18.42", "sey18.66b","hol18.39d")))

# remove samples with low read count
rbcl_rel_abund %>% 
  group_by(sample_id, protocol) %>% 
  summarise(sum=sum(n_reads)) %>% 
  arrange(sum)
# remove failed sample lib18.01
rbcl_rel_abund <- rbcl_rel_abund %>% 
  filter(sample_id != "lib18.01" )

# remove outlier sey18.22a, likely cross contam
rbcl_rel_abund <- rbcl_rel_abund %>% 
  filter(sample_id != "sey18.22a")

# now remove rows that contain 0s
rbcl_rel_abund <- rbcl_rel_abund %>% 
  filter(n_reads!=0)

rbcl_rel_abund <- rbcl_rel_abund %>% 
  select(-protocol)

rbcl_rel_abund
```

Wrangle rbcL taxonomy table
```{r taxonomy}
idtax <- idtax %>% 
  dplyr::rename(domain=V2, kingdom=V3, phylum=V4,class=V5 ,order=V6, family=V7, genus=V8, species=V9) %>% 
  dplyr::rename(domainConf=V21,kingdomConf=V31, phylumConf=V41, classConf = V51, orderConf=V61, familyConf=V71, genusConf=V81, speciesConf=V91) %>% 
  mutate_if(is.character, funs(replace(., str_detect(., "unclassified"), NA_character_))) %>% 
  mutate(id_level = case_when(is.na(domain) ~"root",
                              is.na(kingdom) ~"domain",
                              is.na(phylum)~"kingdom",
                              is.na(class)~"phylum",
                              is.na(order)~"class",
                              is.na(family)~"order",
                              is.na(genus)~"family",
                              is.na(species)~"genus",
                              TRUE ~ "species")) %>% 
  mutate(assignment = case_when(id_level =="root"~"root",
                                id_level == "domain"~ domain,
                                id_level == "kingdom"~kingdom,
                                id_level == "phylum"~phylum,
                                id_level == "class"~class,
                                id_level == "order"~order,
                                id_level == "family"~family,
                                id_level == "genus"~genus,
                                id_level == "species"~species))
idtax %>% nrow() # 652 ASVs
# remove ASVs not found in snow samples
idtax <- idtax %>% 
  filter(asv %in% rbcl_rel_abund$asv)
idtax %>% nrow()
```

For convenience, assign each rbcL ASV a numerical ID
```{r asv_key}
# make rbcl asv id key
rbcl_asv_id_key <- idtax %>% 
  select(asv) %>% 
  rownames_to_column("asv_id") %>% 
  mutate(asv_id = str_pad(asv_id,3,pad="0")) %>% 
  arrange(asv_id)

# add in asv_id to asv tbl
rbcl_rel_abund <- rbcl_asv_id_key %>% 
  right_join(rbcl_rel_abund, by="asv") %>% 
  arrange(asv_id) %>% 
  select(-asv)

# add asv_id to taxonomy tbl
idtax <- rbcl_asv_id_key %>% 
  right_join(idtax, by="asv") %>% 
  select(-asv) %>% 
  arrange(asv_id)

idtax %>% distinct(genus)
```

Manually change some of the taxonomy that was incorrectly labelled
```{r}
idtax <- idtax %>% 
  mutate(genus = case_when(str_detect(genus,"Stichococcus")~"Stichococcus",
                           str_detect(genus,"Myrmecia")~"Myrmecia",
                           str_detect(genus, "Dicty")~"Dictyochloropsis",
                           str_detect(genus,"Trebouxia")~"Trebouxia",
                           str_detect(family,"Stichococcus")~"Stichococcus", # a mis-alignment of columns in the ref db for this one
                           TRUE~genus),
         species = case_when(str_detect(genus,"Stichococcus")~NA_character_, # not named species, removing
                           TRUE~species),
         id_level = case_when(str_detect(genus,"Stichococcus")~"genus", # and update this
                              TRUE~id_level))
# species list
idtax %>% distinct(species)
```

# 18S ---------------------------

Wrangle 18S rel abundance tables
```{r 18S_wrangle}
unfilt_18s_ra_2019 <- asv.tbl.18s %>%
  t() %>% 
  as_tibble(rownames = "asv")

# change column names
colnames(unfilt_18s_ra_2019)[-1] <- colnames(unfilt_18s_ra_2019)[-1] %>% 
  str_remove("_S.*_") %>% # remove _S*_ ending
  str_replace_all("-",".") #replace hyphen with period

# remove duplocate protocols, remove p from name
colnames(unfilt_18s_ra_2019) # look for dups
unfilt_18s_ra_2019 <- unfilt_18s_ra_2019 %>% 
  select(-gar18.01.p2, -sax18.01.p1, -sey18.25.p1, -sey18.74.p1, -sky18.12.p1)

# now remove p from sample Id
colnames(unfilt_18s_ra_2019)[-1] <- colnames(unfilt_18s_ra_2019)[-1] %>% 
  str_remove("\\.p[:digit:]")

# tidy long form, remove remaining samples not in rbcl data set (water, etc)
s18_rel_abund <- unfilt_18s_ra_2019 %>% 
  gather(key = "sample_id", val = "n_reads", -asv) %>% 
  filter(sample_id %in% rbcl_rel_abund$sample_id)

# add percent asv
s18_rel_abund <- s18_rel_abund %>% 
  group_by(sample_id) %>% 
  mutate(rel_abund = n_reads/sum(n_reads)) %>% 
  ungroup()

# n reads per sample
s18_rel_abund %>% 
  group_by(sample_id) %>% 
  summarise(sum_n = sum(n_reads)) %>% 
  arrange(sum_n) # max 76k, min 21k reads per sample

s18_rel_abund <- s18_rel_abund %>% 
  filter(sample_id != "gar18.04") %>% # remove gar4, no reads (?)
  filter(rel_abund != 0) # remove 0s
s18_rel_abund
```

Wrangle 18S taxonomy
```{r}
s18_taxa_silva <- s18_taxa_silva %>% 
  dplyr::rename(domain=V2,phylum=V3, class=V4 ,order=V5, family=V6, genus=V7, domainConf=V21,phylumConf=V31, classConf=V41, orderConf = V51, familyConf=V61 ,genusConf=V71) %>% 
  mutate_if(is.character, funs(replace(., str_detect(., "unclassified"), NA_character_))) %>% 
  mutate(id_level = case_when(is.na(domain) ~"none",
                              is.na(phylum) ~"domain",
                              is.na(class)~"phylum",
                              is.na(order)~"class",
                              is.na(family)~"order",
                              is.na(genus)~"family",
                              TRUE ~ "genus")) %>% 
  mutate(assignment = case_when(id_level == "none"~ "none",
                                id_level == "domain"~ domain,
                                id_level == "phylum"~phylum,
                                id_level == "class"~class,
                                id_level == "order"~order,
                                id_level == "family"~family,
                                id_level == "genus"~genus,
                                TRUE ~ "ERROR"))
s18_taxa_silva
##### some weird additions to the 18s taxonomy tbl from SILVA, _fa, _ge, only in some cases...
```

For convenience, assign each 18S ASV a numerical ID
```{r}
# make rbcl asv id key
s18_asv_id_key <- s18_taxa_silva %>% 
  select(asv) %>% 
  rownames_to_column("asv_id") %>% 
  mutate(asv_id = str_pad(asv_id,3,pad="0")) %>% 
  arrange(asv_id)

# add in asv_id to asv tbl
s18_rel_abund <- s18_asv_id_key %>% 
  right_join(s18_rel_abund, by="asv") %>% # right_join just a shortcut so that asv_id is first column
  arrange(asv_id) %>% 
  select(-asv)

# add asv_id to taxonomy tbl
s18_taxa_silva <- s18_asv_id_key %>% 
  right_join(s18_taxa_silva, by="asv") %>% 
  select(-asv) %>% 
  arrange(asv_id)

```

Wrangle field data
```{r field}

field <- field %>%
  mutate(
    # make new col with 3 letter mountain code
    mountain = str_sub(sample_id, 1,3), 
    # lump some of the color snow categories
    color_snow = case_when(color_snow %in% c("white","yellow") ~ "other",
                           color_snow %in% c("purple","red") ~ "pink", 
                           is.na(color_snow) ~ "other",
                           TRUE ~ color_snow),
    # change NA habitat to other
    habitat = if_else(is.na(habitat), "other", habitat)
  )
```

```{r output}
rbcl_rel_abund %>% write_csv(here("data","clean","rbcl_rel_abund_clean.csv"))
idtax %>% write_csv(here("data","clean","rbcl_taxonomy_clean.csv"))
rbcl_asv_id_key %>% write_csv(here("data","clean","rbcl_asv_key.csv"))

write_csv(field, here("data","clean","field_clean.csv"))
```