### 2019-11-02 manuscript/figures.Rmd
init

---
title: "Figures"
output: pdf_document
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=12, fig.height=8, fig.path='figs/', echo=FALSE, warning=FALSE, message=FALSE)
```

```{r input, include=F}
# field <- knitr::include_graphics("temp.png")
asvs <- include_graphics("figs/2_taxonomy/combined.png")
communities <- include_graphics("figs/3_community/combined.png")
# depth
```

```{r pressure, echo=FALSE, fig.cap="A caption", out.width = '100%'}
communities
```

Figure 1.