### 2019-11-02 manuscript/supplemental.Rmd
init

---
title: "Supplemental Figures"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r table}
knitr::kable(mtcars[1:5,, 1:5], caption = "A table caption")
```