### 2019-11-02 code/depth_comparison.Rmd
ordered otus

---
title: "Depth sample comparison"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)
```

```{r include=F}
rbcl_rel_abund <- read_csv(here("data/03_filtered/rbcl/depth_rbcl_rel_abund.csv"))
s18_all_rel_abund <- read_csv(here("data/03_filtered/18s/depth_18s_rel_abund.csv"))

rbcl_tax <- read_csv(here("data/02_tidied/tidied.rbcl_taxonomy.csv"))
s18_tax <- read_csv(here("data/02_tidied/tidied_silva_18s_assignments.csv"))

rbcl_otus <- read_csv(here("output/asv_otu_key.csv"))

field <- read_csv(here("data/02_tidied/tidied_field.csv"))
```

```{r rbcl_wrangle, include=F}
rbcl <- rbcl_rel_abund %>% 
  left_join(rbcl_tax, by="asv_id") %>% 
  left_join(field, by="sample_id") %>% 
  left_join(rbcl_otus, by="asv_id")
glimpse(rbcl)
rbcl %>% distinct(sample_id, alias, depth_cm)

depth_sample_info <- rbcl %>% 
  distinct(sample_id, alias, depth_cm) %>% 
  mutate(depth_chr = if_else(is.na(depth_cm), "surface", "sub-surface"),
         top_sample = case_when(alias == "H2.10"~"H2.10*", # add a star, same site, but different date
                                alias == "H3.4"~"H2.10*",
                                alias == "H4.3"~"H4.4",
                                alias == "S9.2"~"S9.1",
                                TRUE ~ alias)) %>% 
  select(sample_id, top_sample, depth_chr)
depth_sample_info

rbcl_1 <- rbcl %>% 
  left_join(depth_sample_info, by="sample_id") %>% 
  group_by(top_sample, depth_chr, otu_name) %>% 
  summarise(rel_abund= sum(rel_abund)) %>% 
  ungroup()

rbcl_1
```

```{r}

otu_ord <- c("Chlainomonas A","Sanguina B", "Other Chloromonas","Chloromonas krienitzii","Chloromonas muramotoi C","Chloromonas D", "Algae E","Chloromonas F", "Trebouxiophyceae G")

depth <- ggplot(rbcl_1, aes(x=top_sample, y=rel_abund, fill = otu_name %>% fct_relevel(otu_ord)))+
  geom_bar(stat="identity") +
  facet_grid(rows = vars(depth_chr %>% fct_relevel("surface","sub-surface"))) +
  labs(x="Sample ID", y="Relative abundance", fill="OTU", tag="D")
depth
```

```{r 18s_wrangle, include=F}
s18 <- s18_all_rel_abund %>% 
  left_join(s18_tax, by="asv_id") %>% 
  left_join(field, by="sample_id") %>% 
  left_join(depth_sample_info, by="sample_id") %>% 
  select(top_sample, depth_chr, phylum:genus, rel_abund)

glimpse(s18)
```

```{r}
ggplot(s18, aes(x=top_sample, y=rel_abund, fill = phylum))+
  geom_bar(stat="identity") +
  facet_grid(rows = vars(depth_chr %>% fct_relevel("surface","sub-surface"))) +
  labs(x="Sample ID", y="Relative abundance", fill="OTU")
```
More fungi on surface, one sample with more rotifers on surface, more ciliates on surface also

```{r output, include=FALSE}
ggsave(filename = here("manuscript/figs/4_depth_comparison/rbcl.pdf"), plot=depth)
```