### 2019-11-02 manuscript/supplemental.Rmd
describe each sup fig

---
title: "Supplemental Figures"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Supplemental figures:
 S1) Map of sample locations
 S2) Table of sample site information
 S3) Photos of field sites
 S4) Microphotos of snow algae samples
 S5) Plots of snow algae 18S community composition. A. All 18S, Silva assignments. B. Chlorophyta assignments with custom reference database. 
 S6) Maximum likelihood tree of *rbc*L amplicon
 S7) Plot of base differences between Genbank 18S and *rbc*L segments
 S8) Phylogenetic diversity of snow algae. A. Plot of Faith's phylogenetic diversity in snow algae communities by date and elevation B. Linear model results comparing date and elevation as explanatory variables.

```{r table}
knitr::kable(mtcars[1:5,, 1:5], caption = "A table caption")
```