### 2019-11-04 code/2018_sample_map.Rmd
minor, renamed obj

---
title: "Map of 2018 snow algae sample locations"
output: html_notebook
---

```{r libs, include=F}
library(tidyverse)
library(lubridate)
library(ggmap)

library(ggdark)
```

```{r include=F}
field <- read_csv(here::here("data/02_tidied/tidied_field.csv"))
cities <- read_csv(here::here("output/map/city_coords.csv"))

load(here::here("output/map/raw_google_map.rda")) # loads .rda called "algae_map"... not sure how avoid using.rda format, maybe png?
```

```{r all_samples_map, echo=F}
# one point per mountain, with moutain code, colored by mean elevation
# site tbl: each mtn, total samples, visits, n sequenced samples

full_map <- ggmap(algae_map, extent = "device") +
  # jitter points to make it easier to read
  geom_point(data = field, 
              size = 2, alpha=0.5, 
              width = 0.05, height = 0.05, # specify jitter
              aes(color=elev_m)) +
  scale_color_gradientn(colours = terrain.colors(8)[1:7], limits = c(850,2200)) +
  geom_point(data = cities, shape = 15) +
  geom_text(data = cities, aes(label = name), hjust = 1.1)+
  labs(title = "Snow algae sample sites, summer 2018", color="Elevation (m)")

full_map
```

```{r plot_function}
plot_map <- function(my_month){
  # input: chr with capitalized complete month name, eg "May","June"
  # output: ggplot showing the location of samples from that month
  filtered_sample_coords <- field %>% 
    filter( lubridate::month(date, label=TRUE, abbr=FALSE) == my_month)

  ggmap(algae_map, extent = "device") +
    # jitter points to make it easier to read
    geom_jitter(data = filtered_sample_coords, 
                size = 2, alpha=0.7, 
                width = 0.05, height = 0.05, # specify jitter
                aes(color=elev_m)) +
    scale_color_gradientn(colours = terrain.colors(8)[1:7], limits = c(850,2200)) +
    geom_point(data = city_coords, shape = 15) +
    geom_text(data = city_coords, aes(label = name), hjust = 1.1)+
    labs(title = my_month, color="Elevation (m)")}

#test
plot_map("May")
```

```{r save_maps, include=F}

ggsave(filename = here::here("manuscript/figs/map.png"), plot=full_map)

# for the month by month maps
save_map <- function(my_month, num){
  map <- plot_map(my_month)

  ggsave(filename = here::here("figs","map",paste0("map_",num,"_", my_month,".png")),
       plot = map, height = 15, width = 20, units = "cm")
}

save_map("May",1)
save_map("June",2)
save_map("July",3)
save_map("August",4)
save_map("September",5)

```