### 2019-11-04 manuscript/supplemental.Rmd
added 3 figs

---
title: "Supplementary materials"
output: pdf_document
---

```{r,, include=F}
knitr::opts_chunk$set(echo = FALSE)
```

```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(kableExtra)
```

```{r, include=F}
map <- knitr::include_graphics(here::here("manuscript/figs/map.png"))
field <- read_csv(here::here("data/02_tidied/tidied_field.csv"))
field_photos <- knitr::include_graphics(here::here("manuscript/figs/S3_site_field_photos/field_collage.png"))

map_cap <- "S1. Map showing locations of snow algae sampling locations in summer 2018."
table_cap <- "S2. Table of sample site information of sequenced snow algae samples."
field_photo_cap <- "S3. Photos of snow algae field sites."
```

## Figures
 1. Map of sample locations
 2. Photos of field sites
 3. Microphotos of snow algae samples
 4. Plots of snow algae 18S community composition. A. All 18S, Silva assignments. B. Chlorophyta assignments with custom reference database.
 5. Maximum likelihood tree of *rbc*L amplicon
 6. Plot of base differences between Genbank 18S and *rbc*L segments
 7. Phylogenetic diversity of snow algae.

## Tables
 1. Table of sample site information
 2. Software information

```{r, fig.cap=map_cap}
map
```

```{r, fig.cap=field_photo_cap}
field_photos
```

```{r}
field %>% 
   filter(have_seq==T) %>% 
   mutate(notes = if_else(sample_id %in% c("pan18.ro","bre18.rd"), "collected from rock",NA_character_)) %>% 
   select(alias, date, elev_m, lat, lon,color_snow, habitat, notes) %>% 
   arrange(date) %>% 
   kable(caption = table_cap) %>% 
   kable_styling(bootstrap_options = "striped")

```