### 2019-11-05 code/cellcount.Rmd
test

---
title: "Cellct analysis"
output: html_notebook
---

```{r, include=F}
library(tidyverse)
library(here)
library(gridExtra)
test
```

```{r, include=F}
cellct <- read_csv(here("data/02_tidied/tidied.cell_count.csv"))
field <- read_csv(here("data/02_tidied/tidied_field.csv"))
rel_abund <- read_csv(here("data/03_filtered/rbcl/filtered_cleaned_rbcl_rel_abund.csv"))
```

```{r heatmap, echo=F, fig.height=5, fig.width=3}
p <- cellct %>% 
  left_join(field, by="sample_id")

distinct(cellct, sample_id) %>% nrow()

short_names <- p %>% 
  mutate(short_morpho_name = case_when(morpho_sp=="Chlainomonas"~"Cr",
                                       morpho_sp=="Chloromonas cf. brevispina"~"Cb",
                                       morpho_sp=="Chloromonas cf. nivalis"~"Cn",
                                       morpho_sp=="Chloromonas krienitzii"~"Ck",
                                       morpho_sp=="Green cell"~"G",
                                       morpho_sp=="Other"~"O",
                                       morpho_sp=="Sanguina nivaloides"~"Sn"))

my_theme <- theme(axis.text.y=element_blank(),
                  axis.ticks.y=element_blank(), 
                  legend.position = "none")

cellct_heatmap <- ggplot(short_names, aes(y=alias %>% fct_reorder(elev_m), 
              x=short_morpho_name)) +
  geom_tile(aes(alpha = percent_of_sample, fill = morpho_sp)) + # color = "white" %>% fct_reorder(morpho_order)
  labs(x="Cell morphology", y= "Elevation", fill = "Cell morphology", alpha="Relative abundance") +
  scale_alpha(range = c(0.1, 1)) +
  my_theme

cellct_heatmap
```

```{r elevation_bar, echo=F, fig.height=5, fig.width=1}
elevation_ranges <- short_names %>% 
  distinct(elev_m, sample_id) %>% 
  arrange(elev_m)

my_theme <- theme(axis.title.x =element_blank(),
                  axis.text.x=element_blank(),
                  axis.ticks.x=element_blank(),
                  axis.title.y =element_blank(),
                  axis.text.y=element_blank(),
                  axis.ticks.y=element_blank(),
                  panel.grid.major = element_blank(),
                  panel.grid.minor = element_blank(),
                  panel.background = element_blank(),
                  legend.position = "none")

elev <- ggplot(elevation_ranges, aes(x=sample_id %>% fct_reorder(elev_m), y=2200-elev_m, fill=elev_m))+
  geom_bar(width=1, stat = "identity")+
  scale_fill_gradientn(colors = terrain.colors(8)[1:7]) +
  coord_flip() +
  my_theme +
  # labs(y="Elevation (m)")
  annotate(geom="text", x=3, y=500, label="800 m",
              color="black") +
  annotate(geom="text", x=115, y=500, label="2200 m",
              color="black")

elev
```

```{r}
cellct_in_seq <- short_names %>% 
  filter(have_seq==T)

cust_theme <- theme(legend.position = "none",
                    panel.grid.major.x = element_blank(), # remove vertical lines in background
                    panel.grid.minor.x = element_blank(),
                    axis.line=element_blank(), # remove x axis
                    axis.title.x=element_blank(), 
                    axis.text.x=element_blank(),
                    axis.ticks.x=element_blank(),
                    panel.spacing = unit(-0.01, "lines"), # remove space between facet panels
                    # strip.text.x = element_text(angle = 0, hjust=1.0, vjust = 0.8), # rotate facet labels 90
                    strip.background = element_blank()) # remove grey boxes around facet labs

cellct_order <- c("Cr","Sn","Cb","Ck","Cn","G","O")

cellct_in_seq_bar <- ggplot(cellct_in_seq, aes(x=alias %>% fct_reorder(elev_m), y=percent_of_sample)) +
  geom_bar(aes(fill=short_morpho_name), stat="identity") + # classic stacked relative abundance plot
  coord_flip() +
  facet_grid(~short_morpho_name, switch = "x") +
  labs( x="Sample ID", tag="B") +
  cust_theme

cellct_in_seq_bar
```

```{r}
dir.create(here("manuscript/figs/supp/cellct"))
ggsave(filename=here("manuscript/figs/supp/cellct/cellct_heatmap.pdf"), plot=cellct_heatmap, height = 7, width = 3)
ggsave(here("manuscript/figs/supp/cellct/elevations.pdf"), plot=elev, height=7, width=1)
```