### 2019-11-16 code/18s_asv_cluster.Rmd
not using this file, not sure why ites here

---
title: "18S ASV clustering"
output: html_notebook
---

```{r}
library(tidyverse)
library(DECIPHER)

```

```{r}
aligns <- readDNAStringSet(here("data/asv_seq_data/algae_18s_quality_aligns.fasta"))

# taxonomy w genbank (multiple copies of each designation, no editing)
algae_taxonomy <- read_csv(here("output/algae_18s_assignments_snow_algae_genbank.csv"))
```

```{r}
BrowseSeqs(aligns, highlight = T)

dists <-  aligns %>% 
  DECIPHER::DistanceMatrix(type="matrix", correction = "JC")
```

```{r mds}
set.seed(123)
mds_coords <- dists %>% 
  cmdscale() %>% # ok to use euclidian dist, no null values when comparing relatedness
  as_tibble() %>% 
  add_column(asv_id = colnames(dists), .before=1)

mds_coords_w_taxonomy <- mds_coords %>% 
  left_join(algae_taxonomy, by="asv_id")

ggplot(mds_coords_w_taxonomy, aes(x=V1, y=V2))+
  geom_point(aes(color=best_assignment), size=4)
```

```{r}
perplexity <- 30

set.seed(123)
tsne_full_output <- dists %>% 
  Rtsne(pca=F, perplexity = perplexity)

tsne_pts <- tsne_full_output$Y %>% 
  as_tibble()

tsne_pts_taxonomy <- tsne_pts %>% 
  add_column(asv_id = colnames(dists)) %>% # label each point by it's ASV ID
  left_join(taxonomy, by="asv_id")

taxa.order <- c("Chloromonas","Chloromonas_muramotoi","Chloromonas_hohamii","Chloromonas_krienitzii","Chloromonas_Group_B","Raphidonema_nivale","Stichococcus","other Trebouxiophyceae","Chlainomonas", "Chlainomonas_rubra","Sanguina","Sanguina_aurantia","Sanguina_nivaloides" )

ggplot(tsne_pts_taxonomy, aes(x=V1, y=V2)) +
  # plot points not assigned to genus or species
  geom_point(data = filter(tsne_pts_taxonomy, is.na(plot_taxa)), 
             color = "grey", alpha = 0.8) +
  # color points that were assigned to genus or species
  geom_point(data = filter(tsne_pts_taxonomy, !is.na(plot_taxa)), 
             aes(color = plot_taxa %>% fct_relevel(taxa.order)), 
             alpha = 0.8)
```