### 2019-11-16 code/community.Rmd
init, this makes the heatmap that compares 18s rbcl and cellcount. also the cellcount scatterpie.

---
title: "Community comparison"
output: html_notebook
---

```{r libs, include=F}
library(tidyverse)
library(here)

library(GUniFrac)
library(vegan)

library(scatterpie)
library(ggrepel)
library(gridExtra)
library(ggthemes)
library(ggdark)
```

```{r globals, include=F}
# define theme for all heatmaps
my_heat_theme <- theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x=element_blank())
```

```{r input, include=F}
rbcl_rel_abund <- read_csv(here("data/03_filtered/rbcl/filtered_cleaned_rbcl_rel_abund.csv")) # not looking at the depth samples, controls..
rbcl_tax <- read_csv(here("data/02_tidied/tidied.rbcl_taxonomy.csv"))
rbcl_asv_seqs <- read_csv(here("data/asv_seq_data/rbcl_asv_sequence_key.csv"))
rbcl_otus <- read_csv(here("output/asv_otu_key.csv"))
otu_colors <- read_csv(here("output/etc/otu_colors.csv"))

s18_rel_abund <- read_csv(here("data/03_filtered/18s_algae/18s_algae.filtered_cleaned_rel_abund.csv"))
s18_algae_taxonomy <- read_csv(here("output/algae_18s_assignments_snow_algae_genbank.csv"))
# s18_algae_select_taxonomy <- read_csv(here("output/algae_18s_assignments_w_select_snow_algae.csv")) #### this didnt work very well, need to make consensus for each species

field <- read_csv(here("data/02_tidied/tidied_field.csv"))
cellct <- read_csv(here("data/02_tidied/tidied.cell_count.csv"))

rbcl_tree <- read.tree(here("output/tree/rbcl/all_asvs/rbcl_aligns_all.fasta.treefile")) ### this tree should only contain the asvs we found in the "main" sites, not looking at controls, etc
s18_tree <- read.tree(here("output/tree/algae_18s/algae_18s_seqs.fasta.treefile"))
```

```{r cellct_scatter, echo=F}
cellct_scatter_tbl <- cellct %>% 
  left_join(field, by="sample_id") %>% 
  # not showing all cells in cell count, only Sanguina, Chlainomonas, Chloro b and n
  filter(!(morpho_sp %in% c("Other","Green cell")))

cellct_scatter <- ggplot(cellct_scatter_tbl, aes(x=date, y=elev_m)) +
  geom_jitter(aes(color=morpho_sp, size=percent_of_sample), alpha=0.5, width = 10, height=10) +
  scale_size(range=c(0.5, 10)) +
  labs(color="Morphospecies",size="Relative abundance",y="Elevation (m)",x="Date") +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3))) +
  ggthemes::theme_few()
cellct_scatter
```

```{r cellct_scatterpie, echo=F}
cellct_wide <- cellct %>%
  pivot_wider(names_from = morpho_sp, values_from = percent_of_sample, values_fill = list(percent_of_sample=0)) %>% 
  left_join(field, by="sample_id") %>% 
  select(date, elev_m, `Chloromonas krienitzii`:Other) %>% 
  mutate(date=lubridate::yday(date)) %>% # convert to julian date
  rownames_to_column("rowname") %>% 
  as.data.frame()

celltypes <- cellct %>% filter(!is.na(morpho_sp)) %>% distinct(morpho_sp) %>% pull()

base_plot <- ggplot(field, aes(x=date, y=elev_m))

min_date <- min(cellct_wide$date)
max_date <- max(cellct_wide$date)
date_diff <- max_date - min_date
min_elev <- min(cellct_wide$elev_m)
max_elev <- max(cellct_wide$elev_m)
elev_diff <- max_elev - min_elev

cellct_converted <- cellct_wide %>% 
  mutate(elev_m = (elev_m - min_elev)/elev_diff, # express as fraction
         date = (date - min_date)/date_diff) # " "

scatterpie <- ggplot() +
  geom_scatterpie(aes(x=date, y=elev_m, group=elev_m, r=0.04), data=cellct_converted, cols=celltypes, color=NA, alpha=0.6)
scatterpie

```

```{r cellct_heat, echo=F}
# only plot the samples we want to compare
cellct_tbl_w_seq <- cellct %>% 
  filter(sample_id %in% rbcl_rel_abund$sample_id)

# sanity check
samples_with_cellcounts <- cellct_tbl_w_seq %>% distinct(sample_id) %>% pull(sample_id)
setdiff(rbcl_rel_abund$sample_id, samples_with_cellcounts) # samples we are missing cellcts for

# specify plotting order
cellct_ord <- c("Chlainomonas rubra","Sanguina nivaloides","Chloromonas krienitzii","Chloromonas cf. brevispina","Chloromonas cf. nivalis", "Green cell","Other")

# add metadata
cellct_field <- cellct_tbl_w_seq %>% 
  left_join(field, by="sample_id")

# set colors 
cellct_colors <- c("Chlainomonas rubra"="#bebada",
                   "Sanguina nivaloides"="#FCA69C",
                   "Chloromonas krienitzii"="#8dd3c7",
                   "Chloromonas cf. brevispina"="#8dd3c7",
                   "Chloromonas cf. nivalis"="#8dd3c7",
                   "Green cell"="grey70",
                   "Other"="grey70")

# plot
cellct_heatmap <- ggplot(cellct_field, 
                         aes(x=alias %>% fct_reorder(elev_m), 
                             y=morpho_sp %>% fct_relevel(cellct_ord))) +
  geom_tile(aes(alpha = percent_of_sample, fill = morpho_sp %>% fct_relevel(cellct_ord))) +
  labs(y="Cell morphology") +
  scale_alpha(range = c(0.01, 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x=element_blank()) +
  guides(alpha="none",fill="none") +
  my_heat_theme +
  scale_fill_manual(values = cellct_colors)
cellct_heatmap
```

```{r s18_heat_setup, include=F}
# calculate rel abund as % of Chlorophyta reads
s18_algae_rel_abund <- s18_rel_abund %>% 
  group_by(sample_id) %>% 
  mutate(rel_abund = n_reads/sum(n_reads)) %>% 
  ungroup()

# check whats missing from this compared to rbcl
s18_sample_ids <- s18_algae_rel_abund %>% distinct(sample_id) %>% pull()
rbcl_sample_ids <- rbcl_rel_abund %>% distinct(sample_id) %>% pull()
setdiff(rbcl_sample_ids, s18_sample_ids)

# calculate relative abundance per genus
best_assignment_rel_abund <- s18_algae_rel_abund %>% 
  left_join(s18_algae_taxonomy, by="asv_id") %>% 
  group_by(sample_id, best_assignment) %>% 
  summarise(rel_abund = sum(rel_abund)) %>% 
  ungroup()

s18_rel_abund_lumped <- best_assignment_rel_abund %>% 
  mutate(lumped_best_assignment = case_when(best_assignment=="Sanguina_nivaloides"~"Sanguina nivaloides",
                                     best_assignment %>% str_detect("Chloromonas")~"Chloromonas",
                                     # best_assignment == "Chlorogonium"~"Chlorogonium",
                                     best_assignment=="Koliellaceae"~"Trebouxiophyceae",
                                     TRUE~"Other")) %>% 
  select(-best_assignment)

s18_rel_abund_lumped
```

```{r s18_heat_plot, include=F}
# specify plot order
assign_ord <- c("Sanguina nivaloides", "Chloromonas","Trebouxiophyceae","Other")

# add field data
rel_abund_field <- s18_rel_abund_lumped %>% 
  left_join(field, by="sample_id")

# set colors: sanguina, chloromonas
s18_heat_colors <- c("Sanguina nivaloides"="#FCA69C",
                     "Chloromonas" = "#8dd3c7",
                     "Trebouxiophyceae"="#80b1d3",
                     "Other" = "grey70")

# plot
s18_heatmap <- ggplot(rel_abund_field, 
                       aes(y=lumped_best_assignment %>% fct_relevel(assign_ord), 
                           x=alias %>% fct_reorder(elev_m))) +
  geom_tile(aes(alpha=rel_abund, fill=lumped_best_assignment %>% fct_relevel(assign_ord))) +
  labs(y="18S taxonomy") +
  scale_alpha(range = c(0.01, 1)) +
  guides(alpha="none", fill="none") +
  my_heat_theme +
  scale_fill_manual(values = s18_heat_colors)
s18_heatmap
```

```{r rbcl_heat, include=F}
# group by OTU
rbcl_rel_abund <- rbcl_rel_abund %>% 
  left_join(rbcl_otus, by="asv_id") %>% 
  group_by(sample_id, otu_name) %>% 
  summarise(rel_abund = sum(rel_abund)) %>%   
  ungroup()

# add grouping metadata
rbcl_rel_abund_field <- rbcl_rel_abund %>% 
  left_join(field, by="sample_id")

# specify colors
rbcl_otu_colors <- c("Chlainomonas A"="#bebada",
                     "Sanguina B"="#FCA69C",
                     "Other Chloromonas"="#8dd3c7",
                     "Chloromonas krienitzii"="#8dd3c7",
                     "Chloromonas muramotoi C"="#8dd3c7", 
                     "Chloromonas D"="#8dd3c7", 
                     "Chloromonas F"="#8dd3c7", 
                     "Trebouxiophyceae G"="#80b1d3", 
                     "Unknown E"="grey70")

# specify factor order
otu_ord <- c("Chlainomonas A","Sanguina B", "Other Chloromonas","Chloromonas krienitzii","Chloromonas muramotoi C","Chloromonas D","Chloromonas F", "Trebouxiophyceae G","Unknown E")

rbcl_heatmap <- ggplot(rbcl_rel_abund_field, aes(y=otu_name %>% fct_relevel(otu_ord), x=alias %>% fct_reorder(elev_m))) +
  geom_tile(aes(alpha=rel_abund, fill=otu_name %>% fct_relevel(otu_ord))) +
  labs(y="rbcL OTU") +
  scale_alpha(range = c(0.01, 1)) +
  my_heat_theme +
  guides(alpha="none", fill="none") +
  scale_fill_manual(values= rbcl_otu_colors)
rbcl_heatmap
```

```{r combined_heatmap_setup, echo=F, fig.width=7,fig.height=4}
# adjust tbls so they can be combined
cellct_combine <- cellct_tbl_w_seq %>% 
  mutate(marker = "Cell morphology") %>% 
  dplyr::rename(species=morpho_sp, rel_abund=percent_of_sample) %>% 
  select(sample_id, marker, species, rel_abund)

s18_combine <- s18_rel_abund_lumped %>% 
  mutate(marker = "18S") %>% 
  dplyr::rename(species=lumped_best_assignment) %>% 
  select(sample_id, marker, species, rel_abund)

rbcl_combine <- rbcl_rel_abund %>%
  mutate(marker="rbcL") %>% 
  dplyr::rename(species = otu_name) %>% 
  select(sample_id, marker, species, rel_abund)

# combine the tbls into one
combined <- bind_rows(cellct_combine, s18_combine, rbcl_combine) %>% 
  left_join(field, by="sample_id")

# specify factor order
species_ord <- c("Chlainomonas A","Chlainomonas rubra", 
                "Sanguina B", "Sanguina nivaloides",
                "Chloromonas krienitzii","Chloromonas muramotoi C", "Chloromonas D", "Chloromonas cf. brevispina", "Chloromonas cf. nivalis","Chloromonas F",
                "Other Chloromonas", "Chloromonas",
                "Trebouxiophyceae G", "Trebouxiophyceae",
                "Unknown E", "Green cell", "Other")

# specify colors
plot_colors <- c(rbcl_otu_colors,s18_heat_colors, cellct_colors)

# plot
community_heatmap <- ggplot(combined, aes(x=alias %>% fct_reorder(elev_m), 
                                          y = species %>% fct_relevel(species_ord))) + 
  geom_tile(aes(alpha=rel_abund, 
                fill=species %>% fct_relevel(species_ord) )) + 
  facet_grid(rows = vars(marker %>% fct_relevel(c("Cell morphology","18S","rbcL"))), 
             scales="free_y", space="free_y") +
  scale_alpha(range = c(0.01, 1)) +
  guides(alpha="none", fill="none") +
  scale_fill_manual(values = plot_colors) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5),
        axis.title.y=element_blank()) +
  labs(x="Sample ID")

community_heatmap
```

```{r output, echo=F}
ggsave(here::here("manuscript/figs/cellct_scatterpie.pdf"), plot=scatterpie, width=7, height=5)
ggsave(here::here("manuscript/figs/cellct_axis.pdf"), plot=base_plot, width=5, height=5)

ggsave(here("manuscript/figs/main/community/community_ggplot.pdf"), plot=community_heatmap, width=7, height=4)
```