### 2019-11-16 code/rbcl_community_compare.Rmd
updates... only using the NMDS UniFrac and the Diversity scatterplots from this script at this point, the final heatmap migrated into the community.Rmd since im comparing 18s and cellct in that one

---
title: "rbcL community comparison"
output: html_notebook
---

```{r, echo=FALSE}
knitr::opts_chunk$set(fig.path='figs/', echo=FALSE, warning=FALSE, message=FALSE)
```

```{r libs, include=F}
library(tidyverse)
library(here)
library(GUniFrac)
library(vegan)
library(ggrepel)
# library(vegalite)
# library(fossil)
library(picante)
library(gridExtra)

# ggplot themes
library(ggthemes)
library(ggdark)
```

```{r input, include=F}
rel_abund <- read_csv(here("data/03_filtered/rbcl/filtered_cleaned_rbcl_rel_abund.csv")) # not looking at the depth samples, controls, etc
tax <- read_csv(here("data/02_tidied/tidied.rbcl_taxonomy.csv"))
field <- read_csv(here("data/02_tidied/tidied_field.csv"))

asv_seqs <- read_csv(here("data/asv_seq_data/rbcl_asv_sequence_key.csv"))
otus <- read_csv(here("output/asv_otu_key.csv"))

rbcl.tree <- read.tree(here("output/tree/rbcl/all_asvs/rbcl_aligns_all.fasta.treefile")) ### this tree should only contain the asvs we found in the "main" sites, not looking at controls, etc

# the cell counts and 18S rel abund data
cellct_heatmap <- readRDS(here("output/cellct_w_seq_heatmap.Rda"))
s18_heatmap <- readRDS(here("output/s18_heatmap.Rda"))
```

```{r unifrac, include=F}

ra_wide <- rel_abund %>% 
  select(sample_id, asv_id, rel_abund) %>% 
  arrange(asv_id) %>% 
  pivot_wider(names_from = asv_id, values_from = rel_abund, values_fill = list(rel_abund = 0)) %>% 
  column_to_rownames("sample_id")

# root tree with Trebouxiophyceae
root <- tax %>% 
  filter(class=="Trebouxiophyceae") %>% 
  head(1) %>% 
  pull(asv_id)

rbcl.tree <- rbcl.tree %>% root(outgroup = root, resolve.root=T)

# check for differences between asvs in tree and rel abund table (all names must appear in tree)
asvs_rel_abund <- colnames(ra_wide)
asvs_tree <- rbcl.tree$tip.label %>% sort()

length(asvs_rel_abund)
length(asvs_tree)

setdiff(asvs_rel_abund, asvs_tree) # output elts unique to asvs_rel_abund

# run weighted UniFrac
set.seed(123)
unifracs <- ra_wide %>% 
  GUniFrac::GUniFrac(rbcl.tree, alpha=1) 
unifrac <- unifracs$unifracs[, , "d_1"]

# run NMDS ordination on results 
set.seed(123)
nmds <- metaMDS(unifrac, autotransform = F)
```

```{r unifrac_plot, echo=F, fig.width=5, fig.height=4}
# plot
unifrac_plot_data <- nmds %>% 
  scores() %>% 
  as_tibble(rownames = "sample_id") %>% 
  left_join(field, by="sample_id")

# plot unifrac, colored by elevation
uf_plot <- ggplot(unifrac_plot_data, aes(x=NMDS2, y=NMDS1)) +
  geom_point(aes(color=elev_m), size=4) +
  scale_color_gradientn(colours = terrain.colors(8)[1:7]) +
  labs(color="Elevation (m)", tag="A") +
  geom_text_repel(aes(label=alias), size=3, segment.size = 0.3) +
  theme_classic()
uf_plot
```

```{r setup_rel_abund_barplot, include=F}
otu_rel_abund <- rel_abund %>% 
  left_join(otus, by="asv_id") %>% 
  group_by(sample_id, otu_name) %>% 
  summarise(rel_abund = sum(rel_abund)) %>%   
  left_join(field, by="sample_id") %>%
  ungroup()

# otu_rel_abund_round_to_1_percent <- otu_rel_abund %>% 
#   mutate(rel_abund = if_else(rel_abund <0.01, 0.01, rel_abund)) # artificially boost rel abund <1% to 1% so its visible on plot
# 
# # check rel abund sums ~= 1 (will be slightly higher, due to rounding)
# otu_rel_abund_round_to_1_percent %>%
#   group_by(alias) %>%
#   summarise(sum(rel_abund))
```

```{r plot_barplot, echo=F, fig.width=2.9, fig.height=5}
otu_ord <- c("Chlainomonas A","Sanguina B", "Other Chloromonas","Chloromonas krienitzii","Chloromonas muramotoi C","Chloromonas D", "Unknown E","Chloromonas F", "Trebouxiophyceae G")

my_theme <- theme(axis.text.x = element_text(angle = 45, hjust = 1))

otu_rel_abund %>% distinct(sample_id) %>% nrow()

rbcl_heatmap <- ggplot(otu_rel_abund, aes(x=otu_name %>% fct_relevel(otu_ord), y=alias %>% fct_reorder(elev_m))) +
  geom_tile(aes(alpha=rel_abund, fill=otu_name %>% fct_relevel(otu_ord))) +
  labs(x="rbcL OTU", y= "Elevation", fill = "rbcL OTU", alpha="Relative abundance") +
  scale_alpha(range = c(0.01, 1)) +
  my_theme +
  guides(alpha="none", fill="none")
rbcl_heatmap
```

```{r colors, include=F}
# get colors, so they match across plots
g <- ggplot_build(rbcl_heatmap)

colors <- tibble(otu_name = otu_rel_abund$otu_name,
                 color = g$data[[1]]$fill) %>% 
  distinct(otu_name, color) %>% 
  deframe()

length(colors)
colors

```

Now add in the 18s and cellcount data side by side the rbcL OTU relative abundance data
- set colors same
- add in missing cellcount data
- plot left to right (entire width of page): rbcl, cellct, 18s, with rbcl taking up most width
- 
```{r 18s_rel_abund, fig.width=1.7, fig.height=5}
s18_heat <- s18_heatmap +
  scale_fill_manual(values=c("#D39200", "#93AA00", "#FF61C3","#808080"))
s18_heat
```

```{r cellct, fig.width=2.6, fig.height=5}
cellct_heat <- cellct_heatmap +
  scale_fill_manual(values = c("#F8766D","#D39200","#00BA38","#93AA00","#93AA00","#808080","#808080","#808080"))
cellct_heat
```

```{r elev_strip, fig.width=1.5, fig.height=5}
my_theme <- theme_classic() +
  theme(axis.line=element_blank(),
        axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y = element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

elev_strip_tbl <- otu_rel_abund %>% 
  distinct(alias, elev_m) %>% 
  arrange(-elev_m) %>% 
  add_column(x="column_1") %>% 
  mutate(elev_m = elev_m %>% round(digits=-1))
nrow(elev_strip_tbl)

elev_strip <- ggplot(elev_strip_tbl, aes(x=x, y=alias %>% fct_reorder(elev_m))) +
  geom_tile(aes(fill=elev_m)) +
  scale_fill_gradientn(colours = terrain.colors(8)[1:7]) +
  labs(x="Elevation (m)") +
  my_theme 
elev_strip
```

```{r elev_strip2, fig.width=0.5, fig.height=5}
elev_strip2 <- elev_strip +
  guides(fill="none") +
  geom_text(aes(label=elev_m), size=3)
elev_strip2
```

```{r snow_color_strip, fig.width=0.5, fig.height=5}
color_strip_tbl <- otu_rel_abund %>% 
  arrange(-elev_m) %>% 
  distinct(alias, color_snow, elev_m) %>% 
  add_column(x="column_1")

color_strip_tbl %>% distinct(color_snow)

color_strip <- ggplot(color_strip_tbl, aes(x=x, y=alias %>% fct_reorder(elev_m))) +
  geom_tile(aes(fill=color_snow)) +
  my_theme +
  guides(fill="none") +
  scale_fill_manual(values=c(pink="red",other="grey",green="green",orange="orange",rusty="purple"))
color_strip
```

Faith's Phylogenetic Diversity, sum of branch lengths in a sample's phylogenetic tree (does not take relative abundance into account)
```{r faiths, echo=F}
ra_wide_mat <- ra_wide %>% as.matrix()

faiths <- ra_wide_mat %>% 
  picante::pd(rbcl.tree) %>% 
  # attach elevation data
  rownames_to_column("sample_id") %>% 
  left_join(field, by="sample_id")

faith_date_plot <- ggplot(faiths, aes(y=PD, x=date))+
  geom_point(aes(color=elev_m), size=4) +
  geom_smooth(method = "lm")+
  scale_color_gradientn(colours = terrain.colors(8)[1:7]) +
  geom_text_repel(aes(label=alias))+
  labs(title="Phylogenetic diversity of snow algae rbcL throughout the summer of 2018",y="Faith's phylogenetic diversity (PD)",x="Date", color="Elevation (m)")
faith_date_plot
```

```{r, echo=F}
faith_elev_plot <- ggplot(faiths, aes(y=PD, x=elev_m))+
  geom_point() +
  geom_smooth(method = "lm")+
  geom_text_repel(aes(label=alias))+
  labs(y="Faith's phylogenetic diversity (PD)",x="Elevation (m)", tag="B")
faith_elev_plot
```

Shannon and Simpson diversity indices (these take relative abundance into account, but require OTUs)
```{r, include=F}
otu_rel_abund_mat <- otu_rel_abund %>% 
  select(sample_id, otu_name, rel_abund) %>% 
  pivot_wider(names_from = otu_name, values_from = rel_abund, values_fill = list(rel_abund=0)) %>% 
  column_to_rownames("sample_id") %>% 
  as.matrix()

# use ra_wide_mat for asv based diversity
shannon <- otu_rel_abund_mat %>% 
  diversity(index = "shannon") %>% 
  enframe() %>% 
  dplyr::rename(sample_id = name, shannon = value)
simpson <- otu_rel_abund_mat %>% 
  diversity(index = "simpson") %>% 
  enframe() %>% 
  dplyr::rename(sample_id = name, simpson = value)
otu_diversity <- shannon %>% 
  left_join(simpson, by="sample_id") %>% 
  left_join(field, by="sample_id")

# plot diversity of samples
shannon <- ggplot(otu_diversity, aes(x=elev_m, y=shannon))+
  geom_point() +
  geom_text_repel(aes(label=alias)) +
  geom_smooth(method = "lm")+
  labs(y="Shannon diversity", tag="A")+
  theme(axis.title.x = element_blank())
shannon
```

```{r diversity_plots, include=F}
diversity_plots <- gridExtra::grid.arrange(shannon, faith_elev_plot, nrow=2)

```

Chao
```{r}
# otu_rel_abund_mat %>% 
#   mutate
#   # head(1) %>% 
#   as.matrix() %>% 
#   fossil::chao1()
```

```{r lm_faiths_pd, include=FALSE}
lm_out <- lm(PD ~ date + elev_m, data = faiths)
summary(lm_out)

# just elev
lm_out1 <- lm(log(PD) ~ elev_m, data = faiths) %>% # log transform to normalize
  broom::tidy()

elev_slope <- lm_out1 %>% 
  filter(term == "elev_m") %>% 
  pull(estimate)
elev_p <- lm_out1 %>% 
  filter(term == "elev_m") %>% 
  pull(p.value)

# just date
lm_out2 <- lm(PD ~ date, data = faiths) %>% 
  broom::tidy()

date_slope <- lm_out2 %>% 
  filter(term == "date") %>% 
  pull(estimate)
date_p <- lm_out2 %>% 
  filter(term == "date") %>% 
  pull(p.value)

# check resids
hist(resid(lm_out1))

# calculate co-linearity
car::vif(lm_out)
```

```{r output, include=F}

ggsave(here("manuscript/figs/main/community/rbcl.pdf"),plot=rbcl_heatmap, width = 2.9, height = 5)
ggsave(here("manuscript/figs/main/community/18s.pdf"),plot=s18_heat, width = 1.7, height = 5)
ggsave(here("manuscript/figs/main/community/cellct.pdf"),plot=cellct_heat, width = 2.6, height = 5)
ggsave(here("manuscript/figs/main/community/elev.pdf"),plot=elev_strip, width = 1.5, height = 5)
ggsave(here("manuscript/figs/main/community/elev2.pdf"),plot=elev_strip2, width = 0.5, height = 5)
ggsave(here("manuscript/figs/main/community/snow_color.pdf"),plot=color_strip, width = 0.5, height = 5)

ggsave(filename = here("manuscript/figs/main/uf_plot.png"), plot=uf_plot, width=5, height = 4)
ggsave(filename = here("manuscript/figs/diversity.png"), plot = diversity_plots, height = 10, width = 7)
```