### 2019-12-23 code/community.Rmd
stacked barplor with cellct, 18s, rbcL inprogress

---
title: "Community comparison"
output: html_notebook
---

```{r libs, include=F}
library(tidyverse)
library(here)

library(GUniFrac)
library(vegan)

library(scatterpie)
library(ggrepel)
library(grid)
library(egg)
 library(gridExtra)
library(ggthemes)
library(ggdark)
```

```{r globals, include=F}
# define theme for all heatmaps
my_heat_theme <- theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x=element_blank())
theme_set(dark_theme_light())
```

```{r input, include=F}
rbcl_rel_abund <- read_csv(here("data/03_filtered/rbcl/filtered_cleaned_rbcl_rel_abund.csv")) # not looking at the depth samples, controls..
rbcl_tax <- read_csv(here("data/02_tidied/tidied.rbcl_taxonomy.csv"))
rbcl_asv_seqs <- read_csv(here("data/asv_seq_data/rbcl_asv_sequence_key.csv"))
rbcl_otus <- read_csv(here("output/asv_otu_key.csv"))
otu_colors <- read_csv(here("output/etc/otu_colors.csv"))

s18_rel_abund <- read_csv(here("data/03_filtered/18s_algae/18s_algae.filtered_cleaned_rel_abund.csv"))
s18_algae_taxonomy <- read_csv(here("output/algae_18s_assignments_snow_algae_genbank.csv"))
# s18_algae_select_taxonomy <- read_csv(here("output/algae_18s_assignments_w_select_snow_algae.csv")) #### this didnt work very well, need to make consensus for each species

field <- read_csv(here("data/02_tidied/tidied_field.csv"))
cellct <- read_csv(here("data/02_tidied/tidied.cell_count.csv"))

rbcl_tree <- read.tree(here("output/tree/rbcl/all_asvs/rbcl_aligns_all.fasta.treefile")) ### this tree should only contain the asvs we found in the "main" sites, not looking at controls, etc
s18_tree <- read.tree(here("output/tree/algae_18s/algae_18s_seqs.fasta.treefile"))
```

```{r cellct_scatter, echo=F}
cellct_scatter_tbl <- cellct %>% 
  left_join(field, by="sample_id") %>% 
  # not showing all cells in cell count, only Sanguina, Chlainomonas, Chloro b and n
  filter(!(morpho_sp %in% c("Other","Green cell")))

cellct_scatter <- ggplot(cellct_scatter_tbl, aes(x=date, y=elev_m)) +
  geom_jitter(aes(color=morpho_sp, size=percent_of_sample), alpha=0.5, width = 10, height=10) +
  scale_size(range=c(0.5, 10)) +
  labs(color="Morphospecies",size="Relative abundance",y="Elevation (m)",x="Date") +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3))) +
  ggthemes::theme_few()
cellct_scatter
```

```{r cellct_scatterpie, echo=F}
cellct_wide <- cellct %>%
  pivot_wider(names_from = morpho_sp, values_from = percent_of_sample, values_fill = list(percent_of_sample=0)) %>% 
  left_join(field, by="sample_id") %>% 
  select(date, elev_m, `Chloromonas krienitzii`:Other) %>% 
  mutate(date=lubridate::yday(date)) %>% # convert to julian date
  rownames_to_column("rowname") %>% 
  as.data.frame()

celltypes <- cellct %>% filter(!is.na(morpho_sp)) %>% distinct(morpho_sp) %>% pull()

base_plot <- ggplot(field, aes(x=date, y=elev_m))

min_date <- min(cellct_wide$date)
max_date <- max(cellct_wide$date)
date_diff <- max_date - min_date
min_elev <- min(cellct_wide$elev_m)
max_elev <- max(cellct_wide$elev_m)
elev_diff <- max_elev - min_elev

cellct_converted <- cellct_wide %>% 
  mutate(elev_m = (elev_m - min_elev)/elev_diff, # express as fraction
         date = (date - min_date)/date_diff) # " "

scatterpie <- ggplot() +
  geom_scatterpie(aes(x=date, y=elev_m, group=elev_m, r=0.04), data=cellct_converted, cols=celltypes, color=NA, alpha=0.6)
scatterpie

```

```{r cellct_heat, echo=F}
# only plot the samples we want to compare
cellct_tbl_w_seq <- cellct %>% 
  filter(sample_id %in% rbcl_rel_abund$sample_id)

# sanity check
samples_with_cellcounts <- cellct_tbl_w_seq %>% distinct(sample_id) %>% pull(sample_id)
setdiff(rbcl_rel_abund$sample_id, samples_with_cellcounts) # samples we are missing cellcts for

# specify plotting order
cellct_ord <- c("Chlainomonas rubra","Sanguina nivaloides","Chloromonas krienitzii","Chloromonas cf. brevispina","Chloromonas cf. nivalis", "Green cell","Other")

# add metadata
cellct_field <- cellct_tbl_w_seq %>% 
  left_join(field, by="sample_id")

# set colors 
cellct_colors <- c("Chlainomonas rubra"="#bebada",
                   "Sanguina nivaloides"="#FCA69C",
                   "Chloromonas krienitzii"="#8dd3c7",
                   "Chloromonas cf. brevispina"="#8dd3c7",
                   "Chloromonas cf. nivalis"="#8dd3c7",
                   "Green cell"="grey70",
                   "Other"="grey70")

# plot
cellct_heatmap <- ggplot(cellct_field, 
                         aes(x=alias %>% fct_reorder(elev_m), 
                             y=morpho_sp %>% fct_relevel(cellct_ord))) +
  geom_tile(aes(alpha = percent_of_sample, fill = morpho_sp %>% fct_relevel(cellct_ord))) +
  labs(y="Cell morphology") +
  scale_alpha(range = c(0.01, 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x=element_blank()) +
  guides(alpha="none",fill="none") +
  my_heat_theme +
  scale_fill_manual(values = cellct_colors)
cellct_heatmap
```

```{r s18_heat_setup, include=F}
# calculate rel abund as % of Chlorophyta reads
s18_algae_rel_abund <- s18_rel_abund %>% 
  group_by(sample_id) %>% 
  mutate(rel_abund = n_reads/sum(n_reads)) %>% 
  ungroup()

# check whats missing from this compared to rbcl
s18_sample_ids <- s18_algae_rel_abund %>% distinct(sample_id) %>% pull()
rbcl_sample_ids <- rbcl_rel_abund %>% distinct(sample_id) %>% pull()
setdiff(rbcl_sample_ids, s18_sample_ids)

# calculate relative abundance per genus
best_assignment_rel_abund <- s18_algae_rel_abund %>% 
  left_join(s18_algae_taxonomy, by="asv_id") %>% 
  group_by(sample_id, best_assignment) %>% 
  summarise(rel_abund = sum(rel_abund)) %>% 
  ungroup()

s18_rel_abund_lumped <- best_assignment_rel_abund %>% 
  mutate(lumped_best_assignment = case_when(best_assignment=="Sanguina_nivaloides"~"Sanguina nivaloides",
                                     best_assignment %>% str_detect("Chloromonas")~"Chloromonas",
                                     # best_assignment == "Chlorogonium"~"Chlorogonium",
                                     best_assignment=="Koliellaceae"~"Trebouxiophyceae",
                                     TRUE~"Other")) %>% 
  select(-best_assignment)

s18_rel_abund_lumped
```

```{r s18_heat_plot, include=F}
# specify plot order
assign_ord <- c("Sanguina nivaloides", "Chloromonas","Trebouxiophyceae","Other")

# add field data
rel_abund_field <- s18_rel_abund_lumped %>% 
  left_join(field, by="sample_id")

# set colors: sanguina, chloromonas
s18_heat_colors <- c("Sanguina nivaloides"="#FCA69C",
                     "Chloromonas" = "#8dd3c7",
                     "Trebouxiophyceae"="#80b1d3",
                     "Other" = "grey70")

# plot
s18_heatmap <- ggplot(rel_abund_field, 
                       aes(y=lumped_best_assignment %>% fct_relevel(assign_ord), 
                           x=alias %>% fct_reorder(elev_m))) +
  geom_tile(aes(alpha=rel_abund, fill=lumped_best_assignment %>% fct_relevel(assign_ord))) +
  labs(y="18S taxonomy") +
  scale_alpha(range = c(0.01, 1)) +
  guides(alpha="none", fill="none") +
  my_heat_theme +
  scale_fill_manual(values = s18_heat_colors)
s18_heatmap
```

```{r rbcl_heat, include=F}
# group by OTU
rbcl_rel_abund <- rbcl_rel_abund %>% 
  left_join(rbcl_otus, by="asv_id") %>% 
  group_by(sample_id, otu_name) %>% 
  summarise(rel_abund = sum(rel_abund)) %>%   
  ungroup()

# add grouping metadata
rbcl_rel_abund_field <- rbcl_rel_abund %>% 
  left_join(field, by="sample_id")

# specify colors
rbcl_otu_colors <- c("Chlainomonas A"="#bebada",
                     "Sanguina B"="#FCA69C",
                     "Other Chloromonas"="#8dd3c7",
                     "Chloromonas krienitzii"="#8dd3c7",
                     "Chloromonas muramotoi C"="#8dd3c7", 
                     "Chloromonas D"="#8dd3c7", 
                     "Chloromonas F"="#8dd3c7", 
                     "Trebouxiophyceae G"="#80b1d3", 
                     "Unknown E"="grey70")

# specify factor order
otu_ord <- c("Chlainomonas A","Sanguina B", "Other Chloromonas","Chloromonas krienitzii","Chloromonas muramotoi C","Chloromonas D","Chloromonas F", "Trebouxiophyceae G","Unknown E")

rbcl_heatmap <- ggplot(rbcl_rel_abund_field, aes(y=otu_name %>% fct_relevel(otu_ord), x=alias %>% fct_reorder(elev_m))) +
  geom_tile(aes(alpha=rel_abund, fill=otu_name %>% fct_relevel(otu_ord))) +
  labs(y="rbcL OTU") +
  scale_alpha(range = c(0.01, 1)) +
  my_heat_theme +
  guides(alpha="none", fill="none") +
  scale_fill_manual(values= rbcl_otu_colors)
rbcl_heatmap
```

```{r combined_heatmap_setup, echo=F, fig.width=7,fig.height=4}
# adjust tbls so they can be combined
cellct_combine <- cellct_tbl_w_seq %>% 
  mutate(marker = "Cell morphology") %>% 
  dplyr::rename(species=morpho_sp, rel_abund=percent_of_sample) %>% 
  select(sample_id, marker, species, rel_abund)

s18_combine <- s18_rel_abund_lumped %>% 
  mutate(marker = "18S") %>% 
  dplyr::rename(species=lumped_best_assignment) %>% 
  select(sample_id, marker, species, rel_abund)

rbcl_combine <- rbcl_rel_abund %>%
  mutate(marker="rbcL") %>% 
  dplyr::rename(species = otu_name) %>% 
  select(sample_id, marker, species, rel_abund)

# combine the tbls into one
combined <- bind_rows(cellct_combine, s18_combine, rbcl_combine) %>% 
  left_join(field, by="sample_id")

# specify factor order
species_ord <- c("Chlainomonas A"="#984ea3","Chlainomonas rubra"="#984ea3", #purple
                "Sanguina B"="#ff534a", "Sanguina nivaloides"="#ff534a", #red
                "Chloromonas krienitzii"="#ff7f00","Chloromonas muramotoi C"="#173416", "Chloromonas D"="#265725", "Chloromonas cf.  brevispina"="grey10", "Chloromonas cf. nivalis"="grey10","Chloromonas F"="#b25800", #greens
                "Other Chloromonas"="#4daf4a", "Chloromonas"="#4daf4a", #light greens
                "Trebouxiophyceae G"="#377eb8", "Trebouxiophyceae"="#377eb8", # blue
                "Unknown E"="grey10", "Green cell"="grey10", "Other"="grey70") #greys

# plot
community_heatmap <- ggplot(combined, aes(x=alias %>% fct_reorder(elev_m), 
                                          y = species %>% fct_relevel(names(species_ord)))) + 
  geom_tile(aes(alpha=rel_abund, 
                fill=species %>% fct_relevel(names(species_ord) ))) + 
  facet_grid(rows = vars(marker %>% fct_relevel(c("Cell morphology","18S","rbcL"))), 
             scales="free_y", space="free_y") +
  scale_alpha(range = c(0.01, 1)) +
  guides(alpha="none", fill="none") +
  scale_fill_manual(values = species_ord) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5),
        axis.title.y=element_blank()) +
  labs(x="Sample ID")

community_heatmap
```

```{r rotated-stacked-bar, echo=F}
# specify factor order, so Chlainomonas always on same side, etc
species_colors <- c("Chlainomonas A"="#984ea3","Chlainomonas rubra"="#984ea3", #purple
                "Sanguina B"="#ff534a", "Sanguina nivaloides"="#ff534a", #red
                "Chloromonas krienitzii"="#ff7f00","Chloromonas muramotoi C"="#173416", "Chloromonas D"="#265725", "Chloromonas cf. brevispina"="forestgreen", "Chloromonas cf. nivalis"="lightgreen","Chloromonas F"="#b25800", #greens
                "Other Chloromonas"="#4daf4a", "Chloromonas"="#4daf4a", #light greens
                "Trebouxiophyceae G"="#377eb8", "Trebouxiophyceae"="#377eb8", # blue
                "Unknown E"="grey40", "Green cell"="#b2d8b2", "Other"="grey70") #greys
# backwards
species_colors <- species_colors %>% rev()

# plot
stacked_bar <- ggplot(combined, aes(x=alias %>% fct_reorder(elev_m), y=rel_abund)) +
  geom_bar(aes(fill=species %>% fct_relevel(names(species_colors))), position="stack", stat="identity") +
  coord_flip() +
  facet_grid(cols = vars(marker %>% fct_relevel(c("Cell morphology","18S","rbcL"))), 
             scales="free_y", space="free_y") +
  theme_classic() +
  labs(x="Sample ID", y="Relative abundance")+
  scale_fill_manual(values = species_colors)

stacked_bar +
  theme(legend.position="none")
```

stacked bar legend
```{r}
legend <- stacked_bar %>% cowplot::get_legend()
grid.newpage()
grid.draw(legend)
```

dosent work in rmarkdown
```{r, echo=F}
gt <- ggplot_gtable(ggplot_build(stacked_bar))
gtable_show_layout(gt)

gt$widths[9] = 4*gt$widths[9]
grid::grid.draw(gt)
```

```{r output, echo=F}
ggsave(here::here("manuscript/figs/cellct_scatterpie.pdf"), plot=scatterpie, width=7, height=5)
ggsave(here::here("manuscript/figs/cellct_axis.pdf"), plot=base_plot, width=5, height=5)

# presentation output
ggsave(here("presentation/omics/community_ggplot.pdf"), plot=community_heatmap, width=7, height=4)
```