### 2020-01-11 manuscript/supp/code/field_site_table.Rmd

---
title: "Snow algae sample sites"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=F}
library(tidyverse)
library(kableExtra)
```

```{r, include=F}
field <- read_csv(here::here("data/02_tidied/tidied_field.csv"))

table_data <- field %>% 
  filter(have_seq==T) %>% 
  select(alias, date, lat, lon, elev_m, color_snow, habitat) %>%
  arrange(date) %>% 
  dplyr::rename("Sample ID" = alias, "Date" = date, "Latitude"=lat, "Longitude"=lon, "Elevation (m)"=elev_m, "Color snow"=color_snow, "Habitat"=habitat)
```

```{r, echo=F}
table_data %>% 
  kable(booktabs=T, linesep="") %>% 
  kable_styling()
```