### 2020-01-11 manuscript/supp/code/map/drafts/2018_sample_map.Rmd

---
title: "Map of 2018 snow algae sample locations"
output: html_notebook
---

```{r libs, include=F}
library(tidyverse)
library(lubridate)
library(ggmap)
library(ggrepel)
library(kableExtra)
library(magick)
library(formattable)

library(ggdark)
```

* loads rda from the saved google map so I dont have to re-load it again and again... this could be tidier
* Specify here::here to avoid conflict with lubridate::here
```{r include=F}
field <- read_csv(here::here("data/02_tidied/tidied_field.csv"))
cities <- read_csv(here::here("output/map/city_coords.csv"))
load(here::here("output/map/raw_google_map.rda")) # loads .rda called "algae_map"... not sure how avoid using.rda format, maybe png?
```

```{r}
cities <- cities %>% 
  filter(name == "Vancouver")
```

```{r}
field_summary <- field %>%
  group_by(mountain) %>% 
  drop_na(lat, lon) %>% 
  arrange(date) %>% 
  summarise(n=n(), 
            mean_elev = mean(elev_m), 
            first_visit=min(date) %>% format("%b %d"), 
            last_visit= max(date) %>% format("%b %d"), 
            n_visits= length(unique(date)),
            lat=mean(lat),
            lon=mean(lon),
            date=min(date)) %>% 
  arrange(date) %>% 
  select(-date)
```

```{r all_samples_map, echo=F}
# one point per mountain, with moutain code, colored by mean elevation
# site tbl: each mtn, total samples, visits, n sequenced samples

full_map <- ggmap(algae_map, extent = "device") +
  # jitter points to make it easier to read
  geom_point(data = field, 
              size = 2, alpha=0.5, 
              width = 0.05, height = 0.05, # specify jitter
              aes(color=elev_m)) +
  scale_color_gradientn(colours = terrain.colors(8)[1:7], limits = c(850,2200)) +
  geom_point(data = cities, shape = 15) +
  geom_text(data = cities, aes(label = name), hjust = 1.1)+
  labs(title = "Snow algae sample sites, summer 2018", color="Elevation (m)")

full_map

full_map_2 <-
  ggmap(algae_map, extent = "device") +
  geom_point(data = field_summary, aes(x=lon, y=lat),
              size = 1, alpha=1, 
              color = "red") +
  geom_point(data = cities, shape = 15, color="white") +
  geom_text(data = cities, aes(label = name), hjust = 1.1, color="white") #+
  # geom_text_repel(data=field_summary,
  #                 aes(x=lon, y=lat, label=mountain),
  #                 size=3, vjust=-2, hjust=1) +
  # plot transparent boxes for layers
  # geom_label_repel(data=field_summary,
  #            aes(label=mountain),
  #            label.size = NA,
  #            alpha=0.5, 
  #             label.padding=.1, 
  #             na.rm=TRUE,
  #            seed=123) #+
  #   # plot a second layer exactly the same, the labels for boxes
  # geom_label_repel(data=field_summary,
  #        aes(x=lon, y=lat, label=mountain), 
  #        alpha=1, 
  #        size=3,
  #        segment.color="black",
  #        segment.size = 1, 
  #        seed=123)
 # xlim = c(1, 20)
full_map_2

#missing enchantments point, not on map
```

```{r}
# field_summary_kable <- 
  field_summary %>% 
  mutate(mean_elev = round(mean_elev,0),
         last_visit = if_else(last_visit==first_visit,"",last_visit),
         date_range= if_else(last_visit!="",paste(first_visit, last_visit, sep=" - "), first_visit)) %>% 
  select(mountain, date_range,  mean_elev, n_visits,n) %>% 
  dplyr::rename_all(str_to_title) %>%
  dplyr::rename("Date"=Date_range,"Elevation"=Mean_elev, "N. visits"=N_visits, "Total samples"=N) %>% 
  kable(booktabs=T, format = "html") %>%
  kable_styling("striped",full_width = F) %>%
  row_spec(0:14,color ="white",background ="black") %>% 
  column_spec(1:5, width = "5em")

# generated final via screenshot

  # couldnt get to work
  # tried 
  # https://haozhu233.github.io/kableExtra/save_kable_and_as_image.html
  # https://stackoverflow.com/questions/48834749/error-using-magick-r-to-import-pdf
  # got image_read_pdf to work, still not authorized error for image_read
  # https://stackoverflow.com/questions/42928765/convertnot-authorized-aaaa-error-constitute-c-readimage-453
  # as_image()
  # save_kable(file=paste0(getwd(),"/presentation/omics/field_tbl.pdf"))
  # kable_as_image(filename=paste0(getwd(),"/presentation/omics/field_tbl.pdf"))
field_summary_kable
```

```{r plot_function}
plot_map <- function(my_month){
  # input: chr with capitalized complete month name, eg "May","June"
  # output: ggplot showing the location of samples from that month
  filtered_sample_coords <- field %>% 
    filter( lubridate::month(date, label=TRUE, abbr=FALSE) == my_month)

  ggmap(algae_map, extent = "device") +
    # jitter points to make it easier to read
    geom_jitter(data = filtered_sample_coords, 
                size = 2, alpha=0.7, 
                width = 0.05, height = 0.05, # specify jitter
                aes(color=elev_m)) +
    scale_color_gradientn(colours = terrain.colors(8)[1:7], limits = c(850,2200)) +
    geom_point(data = city_coords, shape = 15) +
    geom_text(data = city_coords, aes(label = name), hjust = 1.1)+
    labs(title = my_month, color="Elevation (m)")}

#test
plot_map("May")
```

```{r save_maps, include=F}

ggsave(filename = here::here("manuscript/figs/map.tiff"), plot=full_map)

# for the month by month maps
# save_map <- function(my_month, num){
#   map <- plot_map(my_month)
#   
#   ggsave(filename = here::here("figs","map",paste0("map_",num,"_", my_month,".png")),
#        plot = map, height = 15, width = 20, units = "cm")
# }
# 
# save_map("May",1)
# save_map("June",2)
# save_map("July",3)
# save_map("August",4)
# save_map("September",5)

# save output for omics presentation
ggsave(filename = here::here("presentation/omics/map.png"))

```