### 2020-01-11 manuscript/supp/code/primer_table.Rmd

---
title: "Primers"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=F}
library(tidyverse)
library(kableExtra)
```

```{r, include=F}
# seq 5->3'
tbl <- tibble(name=c("rbcL369F","rbcL870R", "Euk1181 (Wang et al. 2014)","Euk1624  (Wang et al. 2014)"),
              seq=c("GAACGTGACAAATTAAACAAA","ACCWGAYADACGWAGAGCTT","TTAATTTGACTCAACRCGGG","CGGGCGGTGTGTACAAAGG"))
```

```{r, echo=F}
tbl %>% 
  rename(Name=name, `Sequence (5'-3')`=seq) %>% 
  kable(booktabs=T, linesep="") %>% 
  kable_styling()
```

References:
Wang Y, Tian RM, Gao ZM, Bougouffa S, Qian P-Y. Optimal Eukaryotic 18S and Universal 16S/18S Ribosomal RNA Primers and Their Application in a Study of Symbiosis. PLOS ONE 2014; 9: e90053.