### 2020-01-11 code/exploratory/negative_controls.Rmd
major reorg

---
title: "Negative controls"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)
```

```{r}
neg_controls <- read_csv(here("data/03_filtered/rbcl/water_controls_rbcl_rel_abund.csv"))
taxonomy <- read_csv(here("data/02_tidied/tidied.rbcl_taxonomy.csv"))
otus <- read_csv(here("output/asv_otu_key.csv"))

# source(here("functions/ggplot_functions.R"))  ### not yet working
```

Which batches were these extracted alongside? check notebook

```{r}
joined <- neg_controls %>% 
  left_join(taxonomy, by="asv_id") %>% 
  left_join(otus, by="asv_id")

joined %>% distinct(sample_id)
```

```{r by_otu, echo=F}
otu_rel_abund <- joined %>% 
  group_by(otu_name, sample_id) %>% 
  summarise(rel_abund = sum(rel_abund))

ggplot(otu_rel_abund, aes(x=sample_id, y=rel_abund, fill=otu_name)) +
  geom_bar(stat="identity") +
  labs(x="Sample ID", y="Relative abundance", fill="OTU", tag="D")
```