### 2020-01-13 code/exploratory/green-orange.Rmd
organize

---
title: "Depth comparison Seymour + Hollyburn green -> orange runnels"
output: html_notebook
---
```{r}
library(tidyverse)
library(here)
```

```{r include=F}
rbcl_ra_depth <- read_csv(here::here("data/rel_abund/3_filter/rbcl/depth_rbcl_rel_abund.csv")) # only the "depth samples", see code/wrangle/filter_rbcl_rel_abund.R for selection criteria
rbcl_ra_main <- read_csv(here:::here("data/rel_abund/3_filter/rbcl/rbcl_rel_abund.csv"))
s18_all_rel_abund <- read_csv(here::here("data/rel_abund/3_filter/18s/depth_18s_rel_abund.csv"))
# rbcl_tax <- read_csv(here::here("data/taxonomy/tidied.rbcl_taxonomy.csv"))
s18_tax <- read_csv(here::here("data/taxonomy/tidied_silva_18s_assignments.csv"))
rbcl_otus <- read_csv(here::here("data/otus/otu_rbcl_asvs.csv"))
field <- read_csv(here::here("data/field/tidy_field.csv"))
```

```{r rbcl_wrangle, include=F}
rbcl <- rbcl_ra_depth %>%
  bind_rows(rbcl_ra_main) %>%
  left_join(field, by="sample_id") %>%  
  left_join(rbcl_otus, by="asv_id") %>% 
  filter(alias %in% c("S9.2", "S9.1", "S6.7", "S2.2","H2.10","H3.4")) %>% # choose just six samples
  # excluded H4.3, H4.4 because dominated by Chlainomonas, I want to focus on Chloro krienitzii samples
  mutate(depth = case_when(alias == "S9.2"~"surface", # add new category, whether its surface or subsurface
                           alias == "S9.1"~"depth",
                           alias == "S6.7"~"surface",
                           alias == "S2.2"~"depth",
                           alias == "H2.10"~"surface",
                           alias == "H3.4"~"depth"))

# rbcl %>% distinct(alias, depth, depth_cm)

# select variables of interest, group by otus
rbcl_1 <- rbcl %>% 
  group_by(alias, otu_name, depth) %>% 
  summarise(rel_abund= sum(rel_abund)) %>% 
  ungroup() %>% 
  select(alias, depth, otu_name, rel_abund)

```

```{r}

otu_ord <- c("Chlainomonas A","Sanguina B", "Other Chloromonas","Chloromonas krienitzii","Chloromonas muramotoi C","Chloromonas D", "Algae E","Chloromonas F", "Trebouxiophyceae G")

depth <- ggplot(rbcl_1, aes(x=top_sample, y=rel_abund, fill = otu_name %>% fct_relevel(otu_ord)))+
  geom_bar(stat="identity") +
  facet_grid(rows = vars(depth_chr %>% fct_relevel("surface","sub-surface"))) +
  labs(x="Sample ID", y="Relative abundance", fill="OTU", tag="D")
depth
```

```{r 18s_wrangle, include=F}
s18 <- s18_all_rel_abund %>% 
  left_join(s18_tax, by="asv_id") %>% 
  left_join(field, by="sample_id") %>% 
  left_join(depth_sample_info, by="sample_id") %>% 
  select(top_sample, depth_chr, phylum:genus, rel_abund)

glimpse(s18)
```

```{r}
ggplot(s18, aes(x=top_sample, y=rel_abund, fill = phylum))+
  geom_bar(stat="identity") +
  facet_grid(rows = vars(depth_chr %>% fct_relevel("surface","sub-surface"))) +
  labs(x="Sample ID", y="Relative abundance", fill="OTU")
```
More fungi on surface, one sample with more rotifers on surface, more ciliates on surface also

```{r output, include=FALSE}
#ggsave(filename = here("manuscript/figs/4_depth_comparison/rbcl.pdf"), plot=depth)
```