### 2020-01-22 manuscript/supp/code/diversity.Rmd

---
title: "Diversity"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=F}
library(tidyverse)
library(here)
library(vegan)
library(picante)
library(gridExtra)
library(ape)
library(ggrepel)
library(corrr)
library(lubridate)
```

```{r, include=F}
rbcl_rel_abund <- read_csv(here::here("data/rel_abund/3_filter/rbcl/rbcl_rel_abund.csv")) # not looking at the depth samples, controls, etc
tax <- read_csv(here::here("data/taxonomy/tidied.rbcl_taxonomy.csv"))
field <- read_csv(here::here("data/field/tidy_field.csv"))
asv_seqs <- read_csv(here::here("data/fastas/asv_seq_data/rbcl_asv_sequence_key.csv"))
otus <- read_csv(here::here("data/otus/otu_rbcl_asvs.csv"))
rbcl.tree <- read.tree(here::here("data/tree/rbcl/asv613.fasta.treefile")) ### this tree should only contain the asvs we found in the "main" sites, not looking at controls, etc
```

```{r, include=F}
# get relative abundance of each otu, in wide format, 
otu_rel_abund <- rbcl_rel_abund %>% 
  left_join(otus, by="asv_id") %>% 
  group_by(sample_id, otu_name) %>% 
  summarise(rel_abund = sum(rel_abund)) %>%   
  ungroup() %>% 
  pivot_wider(names_from = otu_name, values_from = rel_abund, values_fill = list(rel_abund=0)) %>% 
  column_to_rownames("sample_id") # vegan::diversity() wants sample identifiers as rownames

# calculate shannon and simpson
shannon <- otu_rel_abund %>% 
  diversity(index = "shannon") %>% 
  enframe() %>% 
  dplyr::rename(sample_id = name, diversity = value) %>% 
  mutate(type="shannon")
# simpson <- otu_rel_abund %>% 
#   diversity(index = "simpson") %>% 
#   enframe() %>% 
#   dplyr::rename(sample_id = name, simpson = value)
# otu_diversity <- shannon %>% 
#   left_join(simpson, by="sample_id") %>% 
#   left_join(field, by="sample_id")

# plot diversity of samples
shannon_plot <- ggplot(shannon %>% left_join(field), aes(x=date, y=diversity))+
  geom_point(aes(color=elev_m), size=3) +
  scale_color_gradientn(colours = terrain.colors(8)[1:7]) +
  geom_text_repel(aes(label=alias)) +
  labs(y="Shannon diversity")+
  theme(axis.title.x = element_blank())
shannon_plot
```

```{r, include=F}
# format
ra_wide <- rbcl_rel_abund %>% 
  select(sample_id, asv_id, rel_abund) %>% 
  arrange(asv_id) %>% 
  pivot_wider(names_from = asv_id, values_from = rel_abund, values_fill = list(rel_abund = 0)) %>% 
  column_to_rownames("sample_id")

ra_wide_mat <- ra_wide %>% as.matrix()

# root tree with Trebouxiophyceae
root <- tax %>% 
  filter(class=="Trebouxiophyceae") %>% 
  head(1) %>% 
  pull(asv_id)

rbcl.tree <- rbcl.tree %>% ape::root(outgroup = root, resolve.root=T)

faiths <- ra_wide_mat %>% 
  picante::pd(rbcl.tree) %>% 
  # attach elevation data
  rownames_to_column("sample_id") %>% 
  rename(diversity=PD) %>% 
  mutate(type="faiths") %>% 
  select(-SR)

faith_plot <- ggplot(faiths %>% left_join(field), aes(y=diversity, x=date))+
  geom_point(aes(color=elev_m), size=3) +
  scale_color_gradientn(colours = terrain.colors(8)[1:7]) +
  geom_text_repel(aes(label=alias))+
  labs(y="Faith's phylogenetic diversity (PD)",x="Date", color="Elevation (m)")

faith_plot
```

```{r diversity_plots, echo=F}
combined <- shannon %>% 
  bind_rows(faiths) %>%
  left_join(field, by="sample_id") %>% 
  mutate(type=str_to_title(type))

ggplot(combined, aes(x=date, y=diversity)) +
  geom_point(aes(color=elev_m), size=3) +
  scale_color_gradientn(colours = terrain.colors(8)[1:7]) +
  geom_text_repel(aes(label=alias), size=2) +
  facet_grid(rows = "type", scales="free_y") +
  labs(y="Diversity", x="Date", color="Elevation (m)")
```

```{r, include=F}
date <- field %>% 
  select(sample_id, date) %>% 
  mutate(date = date %>% lubridate::yday())

cor_df <- faiths %>% 
  rename(faiths = diversity) %>% 
  select(-type) %>% 
  left_join(shannon) %>% 
  rename(shannon = diversity) %>% 
  select(-type) %>% 
  left_join(date)

cor.test(cor_df$faiths, cor_df$date)

cor.test(cor_df$shannon, cor_df$date)

```

**Supplementary figure S6.** Faiths phylogenetic diversity (upper) and Shannon diversity (lower) of snow algae *rbcL* ASVs, plotted against date of sample collection. Samples colored by elevation, and labelled by sample ID. Faith's Pearson's *r*=0.36, p=0.04, Shannon's Pearson's *r*=0.17, p=0.35.